---
title: "Rhizo Assembly Plant Health Analysis"
author: "Abby Sulesky-Grieb"
date: "2023-06-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
```

# Summary
Plants were grown with or without zebularine treatment, labeled as "Zeb" and "Control".
The plants treated with zeb grew visibly slower than the control plants. I harvested plant health measurements of plant height and above ground biomass to track growth progression. Below ground biomass was not collected, as the full root system of each plant was frozen for microbiome analysis. Photos of each root system were taken for analysis with imageJ. 
In the height and biomass data, I will be assessing if the plants harvested by growth stage in zeb and control were the same size, even though they were harvested on different days. I will also be assessing if the plants harvested by time are statistically different due to the different growth rates between zeb and control.

#### Load in data
```{r}
data_rhizo_assembly <- read.csv("/Users/Abby/OneDrive - Michigan State University/Rhizo_assembly/rhizosphere_assembly_plant_data.csv", header=TRUE)

str(data_rhizo_assembly)

# change variables to factors
data_rhizo_assembly$plant_rep <- as.factor(data_rhizo_assembly$plant_rep)
data_rhizo_assembly$group <- as.factor(data_rhizo_assembly$group)
data_rhizo_assembly$series <- as.factor(data_rhizo_assembly$series)
data_rhizo_assembly$stage <- as.factor(data_rhizo_assembly$stage)
data_rhizo_assembly$date <- as.factor(data_rhizo_assembly$date)

```
### Plant height figure
```{r, warning=FALSE}
# code for background theme for ggplot so I don't have to have it in every plot
my_theme <- theme(panel.background = element_rect(fill = "white", colour = "white"), 
             panel.grid.major = element_line(size = 0.25, linetype = 'solid', colour = "light gray"),
             panel.grid.minor = element_line(size = 0.25, linetype = 'solid', colour = "light gray"))

# series name labels
series_labels <- c("growth" = "Growth Stage","time" = "Time Series")

# Colors
colors <- c("#F8766D", "#00bfc4")

# x-axis order
sampling_order <- c("V1", "V2","V3","V4","R1","R4", "R7", "day3", "day7", "day14", "day21", "day35", "day49", "day63")

height_anno <- data.frame(xstar = c(1, 2, 3, 4, 5, 6, 7, 1, 2, 3, 4, 5, 6, 7), 
                          ystar = c(39, 39.5, 39, 39, 39.5, 39, 39, 39, 39, 39, 39, 39, 39, 39),
                   lab = c("**", "ns", "**", "***", "ns", "*", "*", "", "****", "****", "****", "***", "**", "**"),
                   series = c("growth", "growth","growth","growth","growth","growth","growth",
                              "time", "time","time","time","time","time","time"))
height_anno

#  + labs(title="Plant Height by Sampling Series")
height_graph <- ggplot(data_rhizo_assembly, aes(x=factor(stage, level=sampling_order), y=height, fill = group)) +
  geom_boxplot() + my_theme + scale_fill_manual(name = "Treatment Group", values = colors, labels = c("Control", "Zeb")) + xlab("Sample Point") + ylab("Plant Height (cm)") +  stat_summary(fun=mean, colour="black", aes(group=group),
               geom="line", lwd=0.5, lty=1, na.rm=TRUE) + 
  geom_text(inherit.aes=FALSE, data = height_anno, aes(x = xstar,  y = ystar, label = lab), size=4) +
  facet_wrap(~series, scales="free_x", labeller = as_labeller(series_labels)) + theme(axis.text.x=element_blank(), axis.title.x=element_blank())

height_graph 
```

### Plant height stats
```{r, warning=FALSE}
# subset into control v zeb growth and control v zeb time, then by time point
time_data <- subset(data_rhizo_assembly, series == "time")
time_3 <- subset(time_data, stage == "day3")
time_7 <- subset(time_data, stage == "day7")
time_14 <- subset(time_data, stage == "day14")
time_21 <- subset(time_data, stage == "day21")
time_35 <- subset(time_data, stage == "day35")
time_49 <- subset(time_data, stage == "day49")
time_63 <- subset(time_data, stage == "day63")

growth_data <- subset(data_rhizo_assembly, series == "growth")
growth_V1 <- subset(growth_data, stage == "V1")
growth_V2 <- subset(growth_data, stage == "V2")
growth_V3 <- subset(growth_data, stage == "V3")
growth_V4 <- subset(growth_data, stage == "V4")
growth_R1 <- subset(growth_data, stage == "R1")
growth_R4 <- subset(growth_data, stage == "R4")
growth_R7 <- subset(growth_data, stage == "R7")


# compare control vs zeb at time points
height_time_7 <- t.test(height ~ group, data = time_7, na.action = "na.omit")
height_time_7 # ****
height_time_14 <- t.test(height ~ group, data = time_14, na.action = "na.omit")
height_time_14 # ****
height_time_21 <- t.test(height ~ group, data = time_21, na.action = "na.omit")
height_time_21 # ****
height_time_35 <- t.test(height ~ group, data = time_35, na.action = "na.omit")
height_time_35 # *** .0002
height_time_49 <- t.test(height ~ group, data = time_49, na.action = "na.omit")
height_time_49 # ** .001
height_time_63 <- t.test(height ~ group, data = time_63, na.action = "na.omit")
height_time_63 # ** .001

height_growth_V1 <- t.test(height ~ group, data = growth_V1, na.action = "na.omit")
height_growth_V1 # ** .004
height_growth_V2 <- t.test(height ~ group, data = growth_V2, na.action = "na.omit")
height_growth_V2 # .052
height_growth_V3 <- t.test(height ~ group, data = growth_V3, na.action = "na.omit")
height_growth_V3 # ** .007
height_growth_V4 <- t.test(height ~ group, data = growth_V4, na.action = "na.omit")
height_growth_V4 # *** .0008
height_growth_R1 <- t.test(height ~ group, data = growth_R1, na.action = "na.omit")
height_growth_R1 # .21
height_growth_R4 <- t.test(height ~ group, data = growth_R4, na.action = "na.omit")
height_growth_R4 # * .03
height_growth_R7 <- t.test(height ~ group, data = growth_R7, na.action = "na.omit")
height_growth_R7 # * .03






```

### Plant biomass figures
```{r, warning=FALSE}
# Fresh weight
fresh_anno <- data.frame(xstar = c(1, 2, 3, 4, 5, 6, 7, 1, 2, 3, 4, 5, 6, 7), 
                          ystar = c(55, 55, 55, 55, 55, 55, 55, 55, 55, 55, 55, 55, 55, 55),
                   lab = c("", "**", "*", "***", "**", "ns", "*", "", "***", "****", "****", "***", "***", "ns"),
                   series = c("growth", "growth","growth","growth","growth","growth","growth",
                              "time", "time","time","time","time","time","time"))
fresh_anno

# + labs(title="Above-ground Fresh Weight by Sampling Series")
predry_graph <- ggplot(data_rhizo_assembly, aes(x=factor(stage, level=sampling_order), y=predry, fill = group)) +
  geom_boxplot() + my_theme  + scale_fill_manual(name = "Treatment Group", values = colors, labels = c("Control", "Zeb")) + xlab("Sample Point") + ylab("Fresh Biomass (g)") + stat_summary(fun=mean, colour="black", aes(group=group),
               geom="line", lwd=0.5, lty=1, na.rm=TRUE) + facet_wrap(~series, scales="free_x", labeller = as_labeller(series_labels)) + 
  geom_text(inherit.aes=FALSE, data = fresh_anno, aes(x = xstar,  y = ystar, label = lab), size=4) + theme(legend.position="none", axis.text.x=element_blank(), axis.title.x=element_blank())

predry_graph

# Dry weight
dry_anno <- data.frame(xstar = c(1, 2, 3, 4, 5, 6, 7, 1, 2, 3, 4, 5, 6, 7), 
                          ystar = c(15, 15, 15, 15, 15, 15, 15, 15, 15, 15, 15, 15, 15, 15),
                   lab = c("ns", "**", "**", "***", "*", "ns", "ns", "", "ns", "*", "***", "****", "****", "***"),
                   series = c("growth", "growth","growth","growth","growth","growth","growth",
                              "time", "time","time","time","time","time","time"))
dry_anno

#  + labs(title="Above-ground Dry Weight by Sampling Series")
postdry_graph <- ggplot(data_rhizo_assembly, aes(x=factor(stage, level=sampling_order), y=postdry, fill = group)) +
  geom_boxplot() + my_theme + scale_fill_manual(name = "Treatment Group", values = colors, labels = c("Control", "Zeb")) + xlab("Sample Point") + ylab("Dry Biomass (g)") + stat_summary(fun=mean, colour="black", aes(group=group),
               geom="line", lwd=0.5, lty=1, na.rm=TRUE) + facet_wrap(~series, scales="free_x", labeller = as_labeller(series_labels)) + 
  geom_text(inherit.aes=FALSE, data = dry_anno, aes(x = xstar,  y = ystar, label = lab), size=4) + theme(legend.position="none")

postdry_graph

health_combo_graph <- height_graph / predry_graph / postdry_graph + plot_annotation(tag_levels = 'A')
health_combo_graph

ggsave(health_combo_graph, file="/Users/Abby/OneDrive - Michigan State University/Rhizo_assembly/health_multiplot.png", height = 12, width = 9, units = "in")

```
# biomass stats
```{r}
### fresh weight
# compare control vs zeb at time points
fresh_time_7 <- t.test(predry ~ group, data = time_7, na.action = "na.omit")
fresh_time_7 # ***
fresh_time_14 <- t.test(predry ~ group, data = time_14, na.action = "na.omit")
fresh_time_14 # ****
fresh_time_21 <- t.test(predry ~ group, data = time_21, na.action = "na.omit")
fresh_time_21 # ****
fresh_time_35 <- t.test(predry ~ group, data = time_35, na.action = "na.omit")
fresh_time_35 # *** 
fresh_time_49 <- t.test(predry ~ group, data = time_49, na.action = "na.omit")
fresh_time_49 # ***
fresh_time_63 <- t.test(predry ~ group, data = time_63, na.action = "na.omit")
fresh_time_63 # ns .08


fresh_growth_V2 <- t.test(predry ~ group, data = growth_V2, na.action = "na.omit")
fresh_growth_V2 # ** .002
fresh_growth_V3 <- t.test(predry ~ group, data = growth_V3, na.action = "na.omit")
fresh_growth_V3 # * .04
fresh_growth_V4 <- t.test(predry ~ group, data = growth_V4, na.action = "na.omit")
fresh_growth_V4 # *** .0004
fresh_growth_R1 <- t.test(predry ~ group, data = growth_R1, na.action = "na.omit")
fresh_growth_R1 # ** .008
fresh_growth_R4 <- t.test(predry ~ group, data = growth_R4, na.action = "na.omit")
fresh_growth_R4 # ns .08
fresh_growth_R7 <- t.test(predry ~ group, data = growth_R7, na.action = "na.omit")
fresh_growth_R7 # * .02

### dry weight
# compare control vs zeb at time points
dry_time_7 <- t.test(postdry ~ group, data = time_7, na.action = "na.omit")
dry_time_7 # ns
dry_time_14 <- t.test(postdry ~ group, data = time_14, na.action = "na.omit")
dry_time_14 # * .03
dry_time_21 <- t.test(postdry ~ group, data = time_21, na.action = "na.omit")
dry_time_21 # *** .0009
dry_time_35 <- t.test(postdry ~ group, data = time_35, na.action = "na.omit")
dry_time_35 # **** 
dry_time_49 <- t.test(postdry ~ group, data = time_49, na.action = "na.omit")
dry_time_49 # ****
dry_time_63 <- t.test(postdry ~ group, data = time_63, na.action = "na.omit")
dry_time_63 # *** .0001

dry_growth_V1 <- t.test(postdry ~ group, data = growth_V1, na.action = "na.omit")
dry_growth_V1 # ns
dry_growth_V2 <- t.test(postdry ~ group, data = growth_V2, na.action = "na.omit")
dry_growth_V2 # ** .004
dry_growth_V3 <- t.test(postdry ~ group, data = growth_V3, na.action = "na.omit")
dry_growth_V3 # ** .005
dry_growth_V4 <- t.test(postdry ~ group, data = growth_V4, na.action = "na.omit")
dry_growth_V4 # *** .0001
dry_growth_R1 <- t.test(postdry ~ group, data = growth_R1, na.action = "na.omit")
dry_growth_R1 # * .01
dry_growth_R4 <- t.test(postdry ~ group, data = growth_R4, na.action = "na.omit")
dry_growth_R4 # ns 
dry_growth_R7 <- t.test(postdry ~ group, data = growth_R7, na.action = "na.omit")
dry_growth_R7 # ns
```

### Root analysis
```{r}
data_root_length <- read.csv("/Users/Abby/OneDrive - Michigan State University/Rhizo_assembly/Rhizosphere_assembly_Common_Bean/R_Analysis_Files/root_length_data_all.csv", header=TRUE)

str(data_root_length)

data_root_length$series <- as.factor(data_root_length$series)
data_root_length$group <- as.factor(data_root_length$group)
data_root_length$stage <- as.factor(data_root_length$stage)
data_root_length$treatment <- as.factor(data_root_length$treatment)
data_root_length$mean_length_mm <- as.numeric(data_root_length$mean_length_mm)
data_root_length$nodules <- as.numeric(data_root_length$nodules)


root_anno <- data.frame(xstar = c(1, 2, 3, 4, 5, 6, 7, 1, 2, 3, 4, 5, 6, 7), 
                          ystar = c(470, 470, 470, 470, 470, 470, 470, 470, 470, 470, 470, 470, 470, 470),
                   lab = c("****", "****", "***", "ns", "ns", "*", "*", "**", "****", "****", "****", "**", "ns", "**"),
                   series = c("growth", "growth","growth","growth","growth","growth","growth",
                              "time", "time","time","time","time","time","time"))
root_anno

root_length_graph <- ggplot(data_root_length, aes(x=factor(stage, level=sampling_order), y=mean_length_mm, fill = treatment)) + geom_boxplot() + my_theme + scale_fill_manual(name = "Treatment Group", values = colors, labels = c("Control", "Zeb")) + xlab("Sample Point") + ylab("Root Length (mm)") + stat_summary(fun=mean, colour="black", aes(group=group), geom="line", lwd=0.5, lty=1, na.rm=TRUE) + facet_wrap(~series, scales="free_x", labeller = as_labeller(series_labels)) + geom_text(inherit.aes=FALSE, data = root_anno, aes(x = xstar,  y = ystar, label = lab), size=4) + theme(axis.text.x=element_blank(), axis.title.x=element_blank())

root_length_graph


nodule_anno <- data.frame(xstar = c(1, 2, 3, 4, 5, 6, 7, 1, 2, 3, 4, 5, 6, 7), 
                          ystar = c(225, 225, 225, 225, 225, 225, 225, 225, 225, 225, 225, 225, 225, 225),
                   lab = c("ns", "ns", "**", "**", "*", "ns", "ns", "", "ns", "**", "**", "****", "ns", "ns"),
                   series = c("growth", "growth","growth","growth","growth","growth","growth",
                              "time", "time","time","time","time","time","time"))
nodule_anno

nodule_graph <- ggplot(data_root_length, aes(x=factor(stage, level=sampling_order), y=nodules, fill = treatment)) + geom_boxplot() + my_theme + scale_fill_manual(name = "Treatment Group", values = colors, labels = c("Control", "Zeb")) + xlab("Sample Point") + ylab("Nodule Count") + stat_summary(fun=mean, colour="black", aes(group=group), geom="line", lwd=0.5, lty=1, na.rm=TRUE) + facet_wrap(~series, scales="free_x", labeller = as_labeller(series_labels)) + geom_text(inherit.aes=FALSE, data = nodule_anno, aes(x = xstar,  y = ystar, label = lab), size=4) + theme(legend.position="none")

nodule_graph

root_combo_graph <- root_length_graph / nodule_graph + plot_annotation(tag_levels = 'A')
root_combo_graph

ggsave(root_combo_graph, file="/Users/Abby/OneDrive - Michigan State University/Rhizo_assembly/root_multiplot.png", height = 8.5, width = 9, units = "in")
```

### root stats
```{r}
# subset into control v zeb growth and control v zeb time, then by time point
root_time_data <- subset(data_root_length, series == "time")
root_time_3 <- subset(root_time_data, stage == "day3")
root_time_7 <- subset(root_time_data, stage == "day7")
root_time_14 <- subset(root_time_data, stage == "day14")
root_time_21 <- subset(root_time_data, stage == "day21")
root_time_35 <- subset(root_time_data, stage == "day35")
root_time_49 <- subset(root_time_data, stage == "day49")
root_time_63 <- subset(root_time_data, stage == "day63")

root_growth_data <- subset(data_root_length, series == "growth")
root_growth_V1 <- subset(root_growth_data, stage == "V1")
root_growth_V2 <- subset(root_growth_data, stage == "V2")
root_growth_V3 <- subset(root_growth_data, stage == "V3")
root_growth_V4 <- subset(root_growth_data, stage == "V4")
root_growth_R1 <- subset(root_growth_data, stage == "R1")
root_growth_R4 <- subset(root_growth_data, stage == "R4")
root_growth_R7 <- subset(root_growth_data, stage == "R7")

### root length
# compare control vs zeb at time points
root_length_3 <- t.test(mean_length_mm ~ group, data = root_time_3, na.action = "na.omit")
root_length_3 # **
root_length_7 <- t.test(mean_length_mm ~ group, data = root_time_7, na.action = "na.omit")
root_length_7 # ****
root_length_14 <- t.test(mean_length_mm ~ group, data = root_time_14, na.action = "na.omit")
root_length_14 # ****
root_length_21 <- t.test(mean_length_mm ~ group, data = root_time_21, na.action = "na.omit")
root_length_21 # ****
root_length_35 <- t.test(mean_length_mm ~ group, data = root_time_35, na.action = "na.omit")
root_length_35 # ** 
root_length_49 <- t.test(mean_length_mm ~ group, data = root_time_49, na.action = "na.omit")
root_length_49 # ns .06
root_length_63 <- t.test(mean_length_mm ~ group, data = root_time_63, na.action = "na.omit")
root_length_63 # **

root_length_V1 <- t.test(mean_length_mm ~ group, data = root_growth_V1, na.action = "na.omit")
root_length_V1 # ****
root_length_V2 <- t.test(mean_length_mm ~ group, data = root_growth_V2, na.action = "na.omit")
root_length_V2 # ****
root_length_V3 <- t.test(mean_length_mm ~ group, data = root_growth_V3, na.action = "na.omit")
root_length_V3 # ***
root_length_V4 <- t.test(mean_length_mm ~ group, data = root_growth_V4, na.action = "na.omit")
root_length_V4 # ns
root_length_R1 <- t.test(mean_length_mm ~ group, data = root_growth_R1, na.action = "na.omit")
root_length_R1 # ns
root_length_R4 <- t.test(mean_length_mm ~ group, data = root_growth_R4, na.action = "na.omit")
root_length_R4 # *
root_length_R7 <- t.test(mean_length_mm ~ group, data = root_growth_R7, na.action = "na.omit")
root_length_R7 # *


### nodules
nodules_3 <- t.test(nodules ~ group, data = root_time_3, na.action = "na.omit")
nodules_3 # NA
nodules_7 <- t.test(nodules ~ group, data = root_time_7, na.action = "na.omit")
nodules_7 # ns
nodules_14 <- t.test(nodules ~ group, data = root_time_14, na.action = "na.omit")
nodules_14 # **
nodules_21 <- t.test(nodules ~ group, data = root_time_21, na.action = "na.omit")
nodules_21 # **
nodules_35 <- t.test(nodules ~ group, data = root_time_35, na.action = "na.omit")
nodules_35 # ****
nodules_49 <- t.test(nodules ~ group, data = root_time_49, na.action = "na.omit")
nodules_49 # ns
nodules_63 <- t.test(nodules ~ group, data = root_time_63, na.action = "na.omit")
nodules_63 # ns

nodules_V1 <- t.test(nodules ~ group, data = root_growth_V1, na.action = "na.omit")
nodules_V1 # ns .36
nodules_V2 <- t.test(nodules ~ group, data = root_growth_V2, na.action = "na.omit")
nodules_V2 # ns .54
nodules_V3 <- t.test(nodules ~ group, data = root_growth_V3, na.action = "na.omit")
nodules_V3 # **
nodules_V4 <- t.test(nodules ~ group, data = root_growth_V4, na.action = "na.omit")
nodules_V4 # **
nodules_R1 <- t.test(nodules ~ group, data = root_growth_R1, na.action = "na.omit")
nodules_R1 # *
nodules_R4 <- t.test(nodules ~ group, data = root_growth_R4, na.action = "na.omit")
nodules_R4 # ns .68
nodules_R7 <- t.test(nodules ~ group, data = root_growth_R7, na.action = "na.omit")
nodules_R7 # ns .26

```

