# Bean trajectory analysis under Zeb treatment, 16s and ITS component

#### Ludivine Guigard
# based on Samuel Barnett's script
## Loading libraries and main variables
```{r}      
# Libraries for data handling 
library(dplyr) 
library(phyloseq) 
library(decontam)
# Install from GitHub
#if (!requireNamespace("remotes", quietly = TRUE)){
#  install.packages("remotes")
#}
#remotes::install_github("vmikk/metagMisc")
library(metagMisc)  

# Libraries for analysis 
library(vegan) 
library(ape) 
library(ecotraj) 
library(FSA)
library(rcompanion)

# Libraries for plotting 
library(ggplot2)  
library(cowplot)

# Function for extracting legends 
g_legend<-function(a.gplot){   
  tmp <- ggplot_gtable(ggplot_build(a.gplot))   
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")   
  legend <- tmp$grobs[[leg]]   
  return(legend)} 

# Basic plotting theme so as not to continually repeat it 
publication_theme = theme_bw() +   
  theme(axis.text = element_text(size=6),         
        axis.title = element_text(size=7),         
        legend.text = element_text(size=6),         
        legend.title = element_text(size=7, hjust=0.5),         
        strip.text = element_text(size=7),         
        plot.title = element_text(size=8, hjust=0.5))  

present_theme = theme_bw() +   
  theme(axis.text = element_text(size=10),         
        axis.title = element_text(size=12),        
        legend.text = element_text(size=10),         
        legend.title = element_text(size=12, hjust=0.5),         
        strip.text = element_text(size=10),         
        plot.title = element_text(size=14, hjust=0.5))
```

## Loading data

The rarefied data is found in a phyloseq object. Lets take that then divide it into the separate time series and growth stage series.

```{r} 
getwd()
ITS_physeq = readRDS("RA_combined_final.rds")
#View(sample_data(ITS_physeq))
Bac_physeq = readRDS("RA_SAM_16S.rds")
#View(sample_data(Bac_physeq))

ITS_time.rare.physeq = subset_samples(ITS_physeq, series == "time") 
ITS_stage.rare.physeq = subset_samples(ITS_physeq, series == "growth") 

Bac_time.rare.physeq = subset_samples(Bac_physeq, series == "time") 
Bac_stage.rare.physeq = subset_samples(Bac_physeq, series == "growth") 
```
# Time series based analysis

First lets run the trajectory analysis on the time series analysis. We are looking here to see if there are differences in zeb treated and untreated plants and compartment.

## PERMANOVA and Ordination

First run a PERMANOVA to see which factors explain the variation in community composition.

```{r}         
#ITS
perm <- how(nperm = 999) 

ITS_dat.time = data.frame(sample_data(ITS_time.rare.physeq))  
# Get the Bray-Curtis dissimilarity 
ITS_otu.time <- as(otu_table(ITS_time.rare.physeq), "matrix")
ITS_time.BC.dist = vegdist(ITS_otu.time, method = "bray", binary = FALSE, diag = TRUE, upper=TRUE)

# Run adonis2  
set.seed(4242) 
ITS_time.BC.adonis = adonis2(formula = ITS_time.BC.dist ~ treatment*age_days*compartment,
                         permutations = perm, data = ITS_dat.time, by="terms") 
ITS_time.BC.adonis
```
```{r}         
#16S
perm <- how(nperm = 999)
Bac_dat = data.frame(sample_data(Bac_time.rare.physeq))

# Get the Bray-Curtis dissimilarity
Bac_time.BC.dist = vegdist(t(otu_table(Bac_time.rare.physeq)), method="bray", binary=FALSE, diag=TRUE, upper=TRUE)

# Run adonis2 
set.seed(4242)
Bac_time.BC.adonis = adonis2(formula = Bac_time.BC.dist ~ group*stage*compartment, 
                    permutations = perm, data = Bac_dat, by="terms")
Bac_time.BC.adonis
```
As expected, compartment is the primary variable explaining variation in community composition, however development (age_days) and treatment (Zeb treatment) are too. Notably, there are significant interactions as well. Treatment interacts with time (age_days) and time and compartment together, but each explain more variation in composition when taken alone than grouped.
If we select harvestDay instead of age_days, more variability gets explained.

Lets visualize this a bit with ordinations

```{r}         
# Get the PCoA based on Bray-Curtis 
set.seed(4242) 
ITS_time.BC.PCoA = pcoa(ITS_time.BC.dist)  

# Get the axes labels 
ITS_time.BC.PCoA.Xaxis = paste("PCo1 (", round(ITS_time.BC.PCoA$values[1,2]*100, digits=2), "%)", sep="")
ITS_time.BC.PCoA.Yaxis = paste("PCo2 (", round(ITS_time.BC.PCoA$values[2,2]*100, digits=2), "%)", sep="")  
# Convert the PCoA points into a dataframe and add back in the metadata
ITS_time.BC.PCoA.df = data.frame(ITS_time.BC.PCoA$vectors) %>%
  tibble::rownames_to_column(var="SampleID") %>%   
  select(SampleID, Axis.1, Axis.2) %>%
  left_join(data.frame(sample_data(ITS_time.rare.physeq)),              
            by = "SampleID") %>%   
  mutate(day = factor(gsub("day", "", age_days), 
                      levels=c("3", "7", "14", "21", "35", "49", "63"))) 
# PCoA Plot 
ITS_time.BC_PCoA.plot = ggplot(data=ITS_time.BC.PCoA.df, aes(x=Axis.1, y=Axis.2)) + 
  geom_point(aes(fill=treatment, color=treatment, shape=compartment, alpha=day), size=3) +   geom_point(aes(color=treatment, shape=compartment), size=3) +  
  scale_shape_manual(values=c(21, 24)) +   
  scale_fill_manual(values=c("Control" = "black", "Zeb" = "purple")) +  
  scale_color_manual(values=c("Control" = "black", "Zeb" = "purple")) + 
  labs(x=ITS_time.BC.PCoA.Xaxis, y=ITS_time.BC.PCoA.Yaxis) +   
  present_theme
ITS_time.BC_PCoA.plot     
```
```{r}         
# Get the PCoA based on Bray-Curtis
set.seed(4242)
Bac_time.BC.PCoA = pcoa(Bac_time.BC.dist)

# Get the axes labels
Bac_time.BC.PCoA.Xaxis = paste("PCo1 (", round(Bac_time.BC.PCoA$values[1,2]*100, digits=2), "%)", sep="")
Bac_time.BC.PCoA.Yaxis = paste("PCo2 (", round(Bac_time.BC.PCoA$values[2,2]*100, digits=2), "%)", sep="")

# Convert the PCoA points into a dataframe and add back in the metadata
Bac_time.BC.PCoA.df = data.frame(Bac_time.BC.PCoA$vectors) %>%
  tibble::rownames_to_column(var="SampleID") %>%
  select(SampleID, Axis.1, Axis.2) %>%
  left_join(data.frame(sample_data(Bac_time.rare.physeq)), 
            by = "SampleID") %>%
  mutate(day = factor(gsub("day", "", stage), levels=c("3", "7", "14", "21", "35", "49", "63")))

# PCoA Plot
Bac_time.BC_PCoA.plot = ggplot(data=Bac_time.BC.PCoA.df, aes(x=Axis.1, y=Axis.2)) +
  geom_point(aes(fill=group, color=group, shape=compartment, alpha=day), size=3) +
  geom_point(aes(color=group, shape=compartment), size=3) +
  scale_shape_manual(values=c(21, 24)) +
  scale_fill_manual(values=c("Control" = "black", "Zeb" = "purple")) +
  scale_color_manual(values=c("Control" = "black", "Zeb" = "purple")) +
  labs(x=Bac_time.BC.PCoA.Xaxis, y=Bac_time.BC.PCoA.Yaxis) +
  present_theme

Bac_time.BC_PCoA.plot    
```

## Trajectory analysis

Now lets run the trajectory analysis. Since we aren’t following the exact same samples over time, lets try this using a permutative method where we randomly select replicates for each timepoint and run this sampling 100 times and repeat the same analysis for each group.

```{r} 
#Bac
nperm = 100
Bac_time.traj.direct.df = data.frame()
for (i in seq(1:nperm)){
  set.seed(i+42)
  Iter.meta = data.frame(sample_data(Bac_time.rare.physeq)) %>%
    select(SampleID, group, series, stage, plant_rep, compartment) %>%
    mutate(Stage_num = as.numeric(gsub("day", "", stage))) %>%
    group_by(group, series, stage, compartment) %>%
    sample_n(1) %>%
    ungroup %>%
    mutate(PlantID = paste(group, compartment, series, "Iteration", i, sep="_")) %>%
    arrange(PlantID, Stage_num)
  
  D = as.matrix(Bac_time.BC.dist)
  D = D[Iter.meta$SampleID, Iter.meta$SampleID] %>%
    as.dist()
  Plants = Iter.meta$PlantID
  Timepoints = Iter.meta$Stage_num
  
  # Trajectory directionality
  PlantID.traj <- defineTrajectories(D, Plants, Timepoints)

  iter.traj.direct.df = trajectoryMetrics(PlantID.traj) %>%
    rename(PlantID = trajectory) %>%
    left_join(Iter.meta %>%
                select(PlantID, group, series, compartment) %>%
                unique, 
              by = "PlantID") %>%
    mutate(Iteration = i)
  Bac_time.traj.direct.df = rbind(Bac_time.traj.direct.df, iter.traj.direct.df)
}
```

```{r}  
#ITS
# Run the trajectory analysis for permutative sampling of replicates per timepoint.
nperm = 100 
ITS_time.traj.direct.df = data.frame() 
for (i in seq(1:nperm)){   
  set.seed(i+42)   
  Iter.meta = data.frame(sample_data(ITS_time.rare.physeq)) %>%     
    select(SampleID, treatment, series, age_days, plant_rep, compartment) %>%
    mutate(Stage_num = as.numeric(age_days)) %>%     
    group_by(treatment, series, age_days, compartment) %>%     
    sample_n(1) %>%     
    ungroup %>%     
    mutate(PlantID = paste(treatment, compartment, series, "Iteration", i, sep="_")) %>%     
    arrange(PlantID, Stage_num)      
  
  D = as.matrix(ITS_time.BC.dist)   
  D = D[Iter.meta$SampleID, Iter.meta$SampleID] %>%     
    as.dist()   
  Plants = Iter.meta$PlantID   
  Timepoints = Iter.meta$Stage_num      
  
  # Trajectory directionality   
  PlantID.traj <- defineTrajectories(D, Plants, Timepoints)    
  iter.traj.direct.df = trajectoryMetrics(PlantID.traj) %>%     
    dplyr::rename(PlantID = trajectory) %>%     
    left_join(Iter.meta %>%                 
                select(PlantID, treatment, series, compartment) %>%               
                unique,                
              by = "PlantID") %>%     
    mutate(Iteration = i)   
  ITS_time.traj.direct.df = rbind(ITS_time.traj.direct.df, iter.traj.direct.df) 
}
```

First lets take a look at these trajectories across plant compartment within each treatment.

```{r}
# Directionality
Bac_time.comp_dir.wilcox.df = data.frame()
for (grp in c("Control", "Zeb")){
  comp_dir.wilcox = wilcox.test(x=filter(Bac_time.traj.direct.df, group==grp, compartment=="rhizoplane")$directionality,
                                y=filter(Bac_time.traj.direct.df, group==grp, compartment=="rhizosphere")$directionality,
                                conf.int=TRUE, conf.level=0.95)
  Bac_time.comp_dir.wilcox.df = rbind(Bac_time.comp_dir.wilcox.df, 
                             data.frame(group = grp,
                                        ymax = max(filter(Bac_time.traj.direct.df, group==grp)$directionality),
                                        pvalue = comp_dir.wilcox$p.value,
                                        W = comp_dir.wilcox$statistic,
                                        diff = comp_dir.wilcox$estimate))
}

Bac_time.comp_dir.wilcox.df = Bac_time.comp_dir.wilcox.df %>%
  mutate(sig = ifelse(pvalue < 0.001, "p < 0.001",
                      paste("p =", round(pvalue, digits = 3))))

Bac_time.comp_dir.plot = ggplot(data=Bac_time.traj.direct.df, aes(x=compartment, y=directionality)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(height=0, width=0.25, size=2) +
  geom_text(data=Bac_time.comp_dir.wilcox.df, aes(label = sig, y=ymax), size=8*5/14, x=1.5) +
  labs(x = "Plant Compartment", y = "Directionality") +
  present_theme +
  facet_wrap(~group, nrow=1)

Bac_time.comp_dir.plot

#Collapsing both control and Zeb-treated samples: Directionality
Bac_coll_time.comp_dir.wilcox.df = data.frame()
Bac_coll_comp_dir.wilcox = wilcox.test(
  x=filter(Bac_time.traj.direct.df, compartment=="rhizoplane")$directionality,
  y=filter(Bac_time.traj.direct.df, compartment=="rhizosphere")$directionality,
  conf.int=TRUE, conf.level=0.95)  

Bac_coll_time.comp_dir.wilcox.df = rbind(Bac_coll_time.comp_dir.wilcox.df,
                                 data.frame(ymax = max(Bac_time.traj.direct.df$directionality),
                                            pvalue = Bac_coll_comp_dir.wilcox$p.value,
                                            W = Bac_coll_comp_dir.wilcox$statistic,
                                            diff = Bac_coll_comp_dir.wilcox$estimate)) 

Bac_coll_time.comp_dir.wilcox.df = Bac_coll_time.comp_dir.wilcox.df %>%   
  mutate(sig = ifelse(pvalue < 0.001, "p < 0.001",                       
                      paste("p =", round(pvalue, digits = 3))))  

Bac_coll_time.comp_dir.plot = ggplot(data=Bac_time.traj.direct.df, 
                             aes(x=compartment, y=directionality)) +
  geom_boxplot(outlier.shape = NA, fill = c("#E6F9FF","#FFECE6")) +   
  geom_jitter(height=0, width=0.25, size=2) +   
  geom_text(data=Bac_coll_time.comp_dir.wilcox.df, aes(label = sig, y=ymax), size=8*5/14, x=1.5) +   
  labs(x = "Plant Compartment", y = "Directionality") +   
  present_theme 

Bac_coll_time.comp_dir.plot
ggsave("Directionality_Collapsed_Time_16S.png", Bac_coll_time.comp_dir.plot)
```

```{r}    
#ITS
# Directionality 
ITS_time.comp_dir.wilcox.df = data.frame() 
for (grp in c("Control", "Zeb")){   
  comp_dir.wilcox = wilcox.test(
    x=filter(ITS_time.traj.direct.df, treatment==grp, compartment=="rhizoplane")$directionality,                              
    y=filter(ITS_time.traj.direct.df, treatment==grp, compartment=="rhizosphere")$directionality,
    conf.int=TRUE, conf.level=0.95)   
  ITS_time.comp_dir.wilcox.df = rbind(ITS_time.comp_dir.wilcox.df,          
                                  data.frame(treatment = grp,
                                             ymax = max(filter(ITS_time.traj.direct.df, treatment==grp)$directionality),                                        
                                             pvalue = comp_dir.wilcox$p.value,
                                             W = comp_dir.wilcox$statistic,
                                             diff = comp_dir.wilcox$estimate)) 
  }  

ITS_time.comp_dir.wilcox.df = ITS_time.comp_dir.wilcox.df %>%   
  mutate(sig = ifelse(pvalue < 0.001, "p < 0.001",                       
                      paste("p =", round(pvalue, digits = 3))))  

ITS_time.comp_dir.plot = ggplot(data=ITS_time.traj.direct.df, 
                            aes(x=compartment, y=directionality)) +   
  geom_boxplot(outlier.shape = NA) +   
  geom_jitter(height=0, width=0.25, size=2) +   
  geom_text(data=ITS_time.comp_dir.wilcox.df, aes(label = sig, y=ymax), size=8*5/14, x=1.5) +   
  labs(x = "Plant Compartment", y = "Directionality") +   
  present_theme +   
  facet_wrap(~treatment, nrow=1)  

ITS_time.comp_dir.plot

#Collapsing both control and Zeb-treated samples: Directionality
ITS_coll_time.comp_dir.wilcox.df = data.frame()
ITS_coll_comp_dir.wilcox = wilcox.test(
  x=filter(ITS_time.traj.direct.df, compartment=="rhizoplane")$directionality,
  y=filter(ITS_time.traj.direct.df, compartment=="rhizosphere")$directionality,
  conf.int=TRUE, conf.level=0.95)  

ITS_coll_time.comp_dir.wilcox.df = rbind(ITS_coll_time.comp_dir.wilcox.df,
                                 data.frame(ymax = max(ITS_time.traj.direct.df$directionality),
                                            pvalue = ITS_coll_comp_dir.wilcox$p.value,
                                            W = ITS_coll_comp_dir.wilcox$statistic,
                                            diff = ITS_coll_comp_dir.wilcox$estimate)) 

ITS_coll_time.comp_dir.wilcox.df = ITS_coll_time.comp_dir.wilcox.df %>%   
  mutate(sig = ifelse(pvalue < 0.001, "p < 0.001",                       
                      paste("p =", round(pvalue, digits = 3))))  

ITS_coll_time.comp_dir.plot = ggplot(data=ITS_time.traj.direct.df, 
                             aes(x=compartment, y=directionality)) +
  geom_boxplot(outlier.shape = NA, fill = c("#E6F9FF","#FFECE6")) +   
  geom_jitter(height=0, width=0.25, size=2) +   
  geom_text(data=ITS_coll_time.comp_dir.wilcox.df, aes(label = sig, y=ymax), size=8*5/14, x=1.5) +   
  labs(x = "Plant Compartment", y = "Directionality") +   
  present_theme 

ITS_coll_time.comp_dir.plot
ggsave("Directionality_Collapsed_Time_ITS.png", ITS_coll_time.comp_dir.plot)
```
```{r}
# Length 
  #16S
Bac_time.comp_len.wilcox.df = data.frame()
for (grp in c("Control", "Zeb")){
  comp_len.wilcox = wilcox.test(x=filter(Bac_time.traj.direct.df, group==grp, compartment=="rhizoplane")$length,
                                y=filter(Bac_time.traj.direct.df, group==grp, compartment=="rhizosphere")$length,
                                conf.int=TRUE, conf.level=0.95)
  Bac_time.comp_len.wilcox.df = rbind(Bac_time.comp_len.wilcox.df, 
                             data.frame(group = grp,
                                        ymax = max(filter(Bac_time.traj.direct.df, group==grp)$length),
                                        pvalue = comp_len.wilcox$p.value,
                                        W = comp_len.wilcox$statistic,
                                        diff = comp_len.wilcox$estimate))
}

Bac_time.comp_len.wilcox.df = Bac_time.comp_len.wilcox.df %>%
  mutate(sig = ifelse(pvalue < 0.001, "p < 0.001",
                      paste("p =", round(pvalue, digits = 3))))

Bac_time.comp_len.plot = ggplot(data=Bac_time.traj.direct.df, aes(x=compartment, y=length)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(height=0, width=0.25, size=2) +
  geom_text(data=Bac_time.comp_len.wilcox.df, aes(label = sig, y=ymax), size=8*5/14, x=1.5) +
  labs(x = "Plant Compartment", y = "Trajectory length") +
  present_theme +
  facet_wrap(~group, nrow=1)

Bac_time.comp_len.plot

#Collapsing both control and Zeb-treated samples: Length
Bac_coll_time.comp_len.wilcox.df = data.frame() 
Bac_coll_comp_len.wilcox = wilcox.test(x=filter(Bac_time.traj.direct.df, compartment=="rhizoplane")$length,
                              y=filter(Bac_time.traj.direct.df, compartment=="rhizosphere")$length, 
                              conf.int=TRUE, conf.level=0.95)   

Bac_coll_time.comp_len.wilcox.df = rbind(Bac_coll_time.comp_len.wilcox.df,
                                data.frame(ymax = max(Bac_time.traj.direct.df$length),
                                           pvalue = Bac_coll_comp_len.wilcox$p.value,
                                           W = Bac_coll_comp_len.wilcox$statistic,
                                           diff = Bac_coll_comp_len.wilcox$estimate)) 

Bac_coll_time.comp_len.wilcox.df = Bac_coll_time.comp_len.wilcox.df %>%   
  mutate(sig = ifelse(pvalue < 0.001, "p < 0.001",                       
                      paste("p =", round(pvalue, digits = 3))))  

Bac_coll_time.comp_len.plot = ggplot(data=Bac_time.traj.direct.df, aes(x=compartment, y=length)) +
  geom_boxplot(outlier.shape = NA, fill = c("#E6F9FF","#FFECE6")) +   
  geom_jitter(height=0, width=0.25, size=2) +   
  geom_text(data=Bac_coll_time.comp_len.wilcox.df, aes(label = sig, y=ymax), size=8*5/14, x=1.5) +   
  labs(x = "Plant Compartment", y = "Trajectory length") +   
  present_theme 
Bac_coll_time.comp_len.plot
ggsave("Length_Collapsed_Time_16S.png", Bac_coll_time.comp_len.plot)
```
```{r}  
# Length 
  #ITS
ITS_time.comp_len.wilcox.df = data.frame() 
for (grp in c("Control", "Zeb")){   
  comp_len.wilcox = wilcox.test(x=filter(ITS_time.traj.direct.df, treatment==grp,
                                         compartment=="rhizoplane")$length,                                 
                                y=filter(ITS_time.traj.direct.df, treatment==grp, compartment=="rhizosphere")$length,                                 
                                conf.int=TRUE, conf.level=0.95)   
  ITS_time.comp_len.wilcox.df = rbind(ITS_time.comp_len.wilcox.df,                        
                                  data.frame(treatment = grp,                       
                                             ymax = max(filter(ITS_time.traj.direct.df, treatment==grp)$length),                        
                                             pvalue = comp_len.wilcox$p.value,
                                             W = comp_len.wilcox$statistic,
                                             diff = comp_len.wilcox$estimate)) 
  }  

ITS_time.comp_len.wilcox.df = ITS_time.comp_len.wilcox.df %>%   
  mutate(sig = ifelse(pvalue < 0.001, "p < 0.001",                       
                      paste("p =", round(pvalue, digits = 3))))  

ITS_time.comp_len.plot = ggplot(data=ITS_time.traj.direct.df, aes(x=compartment, y=length)) +
  geom_boxplot(outlier.shape = NA) +   
  geom_jitter(height=0, width=0.25, size=2) +   
  geom_text(data=ITS_time.comp_len.wilcox.df, aes(label = sig, y=ymax), size=8*5/14, x=1.5) +   
  labs(x = "Plant Compartment", y = "Trajectory length") +   
  present_theme +   
  facet_wrap(~treatment, nrow=1)  

ITS_time.comp_len.plot

#Collapsing both control and Zeb-treated samples: Directionality
ITS_coll_time.comp_len.wilcox.df = data.frame() 
ITS_coll_comp_len.wilcox = wilcox.test(x=filter(ITS_time.traj.direct.df, compartment=="rhizoplane")$length,
                              y=filter(ITS_time.traj.direct.df, compartment=="rhizosphere")$length, 
                              conf.int=TRUE, conf.level=0.95)   

ITS_coll_time.comp_len.wilcox.df = rbind(ITS_coll_time.comp_len.wilcox.df,
                                data.frame(ymax = max(ITS_time.traj.direct.df$length),
                                           pvalue = ITS_coll_comp_len.wilcox$p.value,
                                           W = ITS_coll_comp_len.wilcox$statistic,
                                           diff = ITS_coll_comp_len.wilcox$estimate)) 

ITS_coll_time.comp_len.wilcox.df = ITS_coll_time.comp_len.wilcox.df %>%   
  mutate(sig = ifelse(pvalue < 0.001, "p < 0.001",                       
                      paste("p =", round(pvalue, digits = 3))))  

ITS_coll_time.comp_len.plot = ggplot(data=ITS_time.traj.direct.df, aes(x=compartment, y=length)) +
  geom_boxplot(outlier.shape = NA, fill = c("#E6F9FF","#FFECE6")) +   
  geom_jitter(height=0, width=0.25, size=2) +   
  geom_text(data=ITS_coll_time.comp_len.wilcox.df, aes(label = sig, y=ymax), size=8*5/14, x=1.5) +   
  labs(x = "Plant Compartment", y = "Trajectory length") +   
  present_theme 
ITS_coll_time.comp_len.plot
ggsave("Length_Collapsed_Time_ITS.png", ITS_coll_time.comp_len.plot)
```

Rhizoplane has significantly more directionality and trajectory length than rhizosphere regardless of the treatment.

# Stage based analysis

Now lets run the trajectory analysis on the stage series analysis. We are looking here to see if there are differences in zeb treated and untreated plants and compartment.

## PERMANOVA and Ordination

First run a PERMANOVA to see which factors explain the variation in community composition.

```{r}         
perm <- how(nperm = 999) 
dat = data.frame(sample_data(ITS_stage.rare.physeq))  

# Get the Bray-Curtis dissimilarity 
ITS_otu.stage <- as(otu_table(ITS_stage.rare.physeq), "matrix")
ITS_stage.BC.dist = vegdist(ITS_otu.stage, method = "bray", binary = FALSE, diag = TRUE, upper=TRUE)

# Run adonis2  
set.seed(4242) 
ITS_stage.BC.adonis = adonis2(formula = ITS_stage.BC.dist ~ treatment*Development*compartment,
                          permutations = perm, data = dat, by="terms") 
ITS_stage.BC.adonis

perm <- how(nperm = 999)
Bac_dat = data.frame(sample_data(Bac_stage.rare.physeq))

# Get the Bray-Curtis dissimilarity
Bac_stage.BC.dist = vegdist(t(otu_table(Bac_stage.rare.physeq)), method="bray", binary=FALSE, diag=TRUE, upper=TRUE)

# Run adonis2 
set.seed(4242)
Bac_stage.BC.adonis = adonis2(formula = Bac_stage.BC.dist ~ group*stage*compartment, 
                    permutations = perm, data = Bac_dat, by="terms")
Bac_stage.BC.adonis
```

As expected, compartment is the primary variable explaining variation in community composition, however stage (development) and group (treatment) are too. We also have the same two interactions in here. This also makes sense since Zeb delays the stages so we might expect more interactions between group and time than group and stage.

Lets visualize this a bit with ordinations

```{r}         
# Get the PCoA based on Bray-Curtis 
set.seed(4242) 
ITS_stage.BC.PCoA = pcoa(ITS_stage.BC.dist)  

# Get the axes labels 
ITS_stage.BC.PCoA.Xaxis = paste("PCo1 (", round(ITS_stage.BC.PCoA$values[1,2]*100, digits=2), "%)", sep="") 
ITS_stage.BC.PCoA.Yaxis = paste("PCo2 (", round(ITS_stage.BC.PCoA$values[2,2]*100, digits=2), "%)", sep="")  

# Convert the PCoA points into a dataframe and add back in the metadata 
ITS_stage.BC.PCoA.df = data.frame(ITS_stage.BC.PCoA$vectors) %>%  
  tibble::rownames_to_column(var="SampleID") %>%   
  select(SampleID, Axis.1, Axis.2) %>%   
  left_join(data.frame(sample_data(ITS_stage.rare.physeq)),              
            by = "SampleID") %>%   
  mutate(stage = factor(Development, levels=c("V1", "V2", "V3", "V4", "R1", "R4", "R7")))  
# PCoA Plot 
ITS_stage.BC_PCoA.plot = ggplot(data=ITS_stage.BC.PCoA.df, aes(x=Axis.1, y=Axis.2)) +   
  geom_point(aes(fill=treatment, color=treatment, shape=compartment, alpha=stage), size=3)+
  geom_point(aes(color=treatment, shape=compartment), size=3) + 
  scale_shape_manual(values=c(21, 24)) +   
  scale_fill_manual(values=c("Control" = "black", "Zeb" = "purple")) +  
  scale_color_manual(values=c("Control" = "black", "Zeb" = "purple")) + 
  labs(x=ITS_stage.BC.PCoA.Xaxis, y=ITS_stage.BC.PCoA.Yaxis) +   
  present_theme  +
  guides(alpha = guide_legend(ncol = 2))
ITS_stage.BC_PCoA.plot

set.seed(4242)
Bac_stage.BC.PCoA = pcoa(Bac_stage.BC.dist)

# Get the axes labels
Bac_stage.BC.PCoA.Xaxis = paste("PCo1 (", round(Bac_stage.BC.PCoA$values[1,2]*100, digits=2), "%)", sep="")
Bac_stage.BC.PCoA.Yaxis = paste("PCo2 (", round(Bac_stage.BC.PCoA$values[2,2]*100, digits=2), "%)", sep="")

# Convert the PCoA points into a dataframe and add back in the metadata
Bac_stage.BC.PCoA.df = data.frame(Bac_stage.BC.PCoA$vectors) %>%
  tibble::rownames_to_column(var="SampleID") %>%
  select(SampleID, Axis.1, Axis.2) %>%
  left_join(data.frame(sample_data(Bac_stage.rare.physeq)), 
            by = "SampleID") %>%
  mutate(stage = factor(stage, levels=c("V1", "V2", "V3", "V4", "R1", "R4", "R7")))

# PCoA Plot
Bac_stage.BC_PCoA.plot = ggplot(data=Bac_stage.BC.PCoA.df, aes(x=Axis.1, y=Axis.2)) +
  geom_point(aes(fill=group, color=group, shape=compartment, alpha=stage), size=3) +
  geom_point(aes(color=group, shape=compartment), size=3) +
  scale_shape_manual(values=c(21, 24)) +
  scale_fill_manual(values=c("Control" = "black", "Zeb" = "purple")) +
  scale_color_manual(values=c("Control" = "black", "Zeb" = "purple")) +
  labs(x=Bac_stage.BC.PCoA.Xaxis, y=Bac_stage.BC.PCoA.Yaxis) +
  present_theme

Bac_stage.BC_PCoA.plot
```

## Trajectory analysis

Now lets run the trajectory analysis. Since we aren’t following the exact same samples over stage, lets try this using a permutative method where we randomly select replicates for each stagepoint and run this sampling 100 stages and repeate the same analysis for each group.

``` {r}    
#ITS
# Run the trajectory analysis for permutative sampling of replicates per stagepoint. 
stage_num.df = data.frame(Stage_num = c(1,2,3,4,5,6,7),                           
                          Development = c("V1", "V2", "V3", "V4", "R1", "R4", "R7"))  
nperm = 100 
ITS_stage.traj.direct.df = data.frame() 
for (i in seq(1:nperm)){   
  set.seed(i+42)   
  Iter.meta = data.frame(sample_data(ITS_stage.rare.physeq)) %>%     
    select(SampleID, treatment, series, Development, plant_rep, compartment) %>%  
    left_join(stage_num.df) %>%     
    group_by(treatment, series, Development, compartment) %>%     
    sample_n(1) %>%     
    ungroup %>%     
    mutate(PlantID = paste(treatment, compartment, series, "Iteration", i, sep="_")) %>%  
    arrange(PlantID, Stage_num)      
  
  D = as.matrix(ITS_stage.BC.dist)   
  D = D[Iter.meta$SampleID, Iter.meta$SampleID] %>%     
    as.dist()   
  Plants = Iter.meta$PlantID   
  stagepoints = Iter.meta$Stage_num      
  
  # Trajectory directionality   
  PlantID.traj <- defineTrajectories(D, Plants, stagepoints)    
  iter.traj.direct.df = trajectoryMetrics(PlantID.traj) %>%     
    rename(PlantID = trajectory) %>%     
    left_join(Iter.meta %>%                 
                select(PlantID, treatment, series, compartment) %>%             
                unique, 
              by = "PlantID") %>%     
    mutate(Iteration = i)   
  
  ITS_stage.traj.direct.df = rbind(ITS_stage.traj.direct.df, iter.traj.direct.df) 
}

#16S
#Run the trajectory analysis for purmutative sampling of replicates per stagepoint.
stage_num.df = data.frame(Stage_num = c(1,2,3,4,5,6,7),
                          stage = c("V1", "V2", "V3", "V4", "R1", "R4", "R7"))

nperm = 100
Bac_stage.traj.direct.df = data.frame()
for (i in seq(1:nperm)){
  set.seed(i+42)
  Iter.meta = data.frame(sample_data(Bac_stage.rare.physeq)) %>%
    select(SampleID, group, series, stage, plant_rep, compartment) %>%
    left_join(stage_num.df) %>%
    group_by(group, series, stage, compartment) %>%
    sample_n(1) %>%
    ungroup %>%
    mutate(PlantID = paste(group, compartment, series, "Iteration", i, sep="_")) %>%
    arrange(PlantID, Stage_num)
  
  D = as.matrix(Bac_stage.BC.dist)
  D = D[Iter.meta$SampleID, Iter.meta$SampleID] %>%
    as.dist()
  Plants = Iter.meta$PlantID
  stagepoints = Iter.meta$Stage_num
  
  # Trajectory directionality
  PlantID.traj <- defineTrajectories(D, Plants, stagepoints)

  iter.traj.direct.df = trajectoryMetrics(PlantID.traj) %>%
    rename(PlantID = trajectory) %>%
    left_join(Iter.meta %>%
                select(PlantID, group, series, compartment) %>%
                unique, 
              by = "PlantID") %>%
    mutate(Iteration = i)
  Bac_stage.traj.direct.df = rbind(Bac_stage.traj.direct.df, iter.traj.direct.df)
}
```
```{r}    
#16S
# Directionality 
Bac_stage.comp_dir.wilcox.df = data.frame()
for (grp in c("Control", "Zeb")){
  comp_dir.wilcox = wilcox.test(x=filter(Bac_stage.traj.direct.df, group==grp, compartment=="rhizoplane")$directionality,
                                y=filter(Bac_stage.traj.direct.df, group==grp, compartment=="rhizosphere")$directionality,
                                conf.int=TRUE, conf.level=0.95)
  Bac_stage.comp_dir.wilcox.df = rbind(Bac_stage.comp_dir.wilcox.df, 
                             data.frame(group = grp,
                                        ymax = max(filter(Bac_stage.traj.direct.df, group==grp)$directionality),
                                        pvalue = comp_dir.wilcox$p.value,
                                        W = comp_dir.wilcox$statistic,
                                        diff = comp_dir.wilcox$estimate))
}

Bac_stage.comp_dir.wilcox.df = Bac_stage.comp_dir.wilcox.df %>%
  mutate(sig = ifelse(pvalue < 0.001, "p < 0.001",
                      paste("p =", round(pvalue, digits = 3))))

Bac_stage.comp_dir.plot = ggplot(data=Bac_stage.traj.direct.df, aes(x=compartment, y=directionality)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(height=0, width=0.25, size=2) +
  geom_text(data=Bac_stage.comp_dir.wilcox.df, aes(label = sig, y=ymax), size=8*5/14, x=1.5) +
  labs(x = "Plant Compartment", y = "Directionality") +
  present_theme +
  facet_wrap(~group, nrow=1)

Bac_stage.comp_dir.plot

#Collapsing both control and Zeb-treated samples: Directionality
Bac_coll_stage.comp_dir.wilcox.df = data.frame()
Bac_coll_stage_comp_dir.wilcox = wilcox.test(
  x=filter(Bac_stage.traj.direct.df, compartment=="rhizoplane")$directionality,
  y=filter(Bac_stage.traj.direct.df, compartment=="rhizosphere")$directionality,
  conf.int=TRUE, conf.level=0.95)  

Bac_coll_stage.comp_dir.wilcox.df = rbind(Bac_coll_stage.comp_dir.wilcox.df,
                                 data.frame(ymax = max(Bac_stage.traj.direct.df$directionality),
                                            pvalue = Bac_coll_stage_comp_dir.wilcox$p.value,
                                            W = Bac_coll_stage_comp_dir.wilcox$statistic,
                                            diff = Bac_coll_stage_comp_dir.wilcox$estimate)) 

Bac_coll_stage.comp_dir.wilcox.df = Bac_coll_stage.comp_dir.wilcox.df %>%   
  mutate(sig = ifelse(pvalue < 0.001, "p < 0.001",                       
                      paste("p =", round(pvalue, digits = 3))))  

Bac_coll_stage.comp_dir.plot = ggplot(data=Bac_stage.traj.direct.df, 
                             aes(x=compartment, y=directionality)) +
  geom_boxplot(outlier.shape = NA, fill = c("#33CCFF","#FF6633")) +   
  geom_jitter(height=0, width=0.25, size=2) +   
  geom_text(data=Bac_coll_stage.comp_dir.wilcox.df, aes(label = sig, y=ymax), size=8*5/14, x=1.5) +
  labs(x = "Plant Compartment", y = "Directionality") +   
  present_theme 

Bac_coll_stage.comp_dir.plot
ggsave("Directionality_Collapsed_Stage_16S.png", Bac_coll_stage.comp_dir.plot)
```

```{r}    
#ITS
# Directionality 
ITS_stage.comp_dir.wilcox.df = data.frame()
for (grp in c("Control", "Zeb")){   
  comp_dir.wilcox = wilcox.test(x=filter(ITS_stage.traj.direct.df, treatment==grp, 
                                         compartment=="rhizoplane")$directionality,
                                y=filter(ITS_stage.traj.direct.df, treatment==grp, 
                                         compartment=="rhizosphere")$directionality,
                                conf.int=TRUE, conf.level=0.95)   
  ITS_stage.comp_dir.wilcox.df = rbind(ITS_stage.comp_dir.wilcox.df,
                                   data.frame(treatment = grp,
                                              ymax = max(filter(ITS_stage.traj.direct.df, treatment==grp)$directionality),
                                              pvalue = comp_dir.wilcox$p.value,
                                              W = comp_dir.wilcox$statistic,
                                              diff = comp_dir.wilcox$estimate)) 
  }  

ITS_stage.comp_dir.wilcox.df = ITS_stage.comp_dir.wilcox.df %>%   
  mutate(sig = ifelse(pvalue < 0.001, "p < 0.001",                       
                      paste("p =", round(pvalue, digits = 3))))  

ITS_stage.comp_dir.plot = ggplot(data=ITS_stage.traj.direct.df, 
                             aes(x=compartment, y=directionality)) +
  geom_boxplot(outlier.shape = NA) +   
  geom_jitter(height=0, width=0.25, size=2) +   
  geom_text(data=ITS_stage.comp_dir.wilcox.df, aes(label = sig, y=ymax), size=8*5/14, x=1.5) +   
  labs(x = "Plant Compartment", y = "Directionality") +   
  present_theme +   
  facet_wrap(~treatment, nrow=1)  

ITS_stage.comp_dir.plot

#Collapsing both control and Zeb-treated samples: Directionality
ITS_coll_stage.comp_dir.wilcox.df = data.frame()
ITS_coll_stage_comp_dir.wilcox = wilcox.test(
  x=filter(ITS_stage.traj.direct.df, compartment=="rhizoplane")$directionality,
  y=filter(ITS_stage.traj.direct.df, compartment=="rhizosphere")$directionality,
  conf.int=TRUE, conf.level=0.95)  

ITS_coll_stage.comp_dir.wilcox.df = rbind(ITS_coll_stage.comp_dir.wilcox.df,
                                 data.frame(ymax = max(ITS_stage.traj.direct.df$directionality),
                                            pvalue = ITS_coll_stage_comp_dir.wilcox$p.value,
                                            W = ITS_coll_stage_comp_dir.wilcox$statistic,
                                            diff = ITS_coll_stage_comp_dir.wilcox$estimate)) 

ITS_coll_stage.comp_dir.wilcox.df = ITS_coll_stage.comp_dir.wilcox.df %>%   
  mutate(sig = ifelse(pvalue < 0.001, "p < 0.001",                       
                      paste("p =", round(pvalue, digits = 3))))  

ITS_coll_stage.comp_dir.plot = ggplot(data=ITS_stage.traj.direct.df, 
                             aes(x=compartment, y=directionality)) +
  geom_boxplot(outlier.shape = NA, fill = c("#33CCFF","#FF6633")) +   
  geom_jitter(height=0, width=0.25, size=2) +   
  geom_text(data=ITS_coll_stage.comp_dir.wilcox.df, aes(label = sig, y=ymax), size=8*5/14, x=1.5) +
  labs(x = "Plant Compartment", y = "Directionality") +   
  present_theme 

ITS_coll_stage.comp_dir.plot
ggsave("Directionality_Collapsed_Stage_ITS.png", ITS_coll_stage.comp_dir.plot)
```

```{r}  
#16S
# Length 
Bac_stage.comp_len.wilcox.df = data.frame()
for (grp in c("Control", "Zeb")){
  comp_len.wilcox = wilcox.test(x=filter(Bac_stage.traj.direct.df, group==grp, compartment=="rhizoplane")$length,
                                y=filter(Bac_stage.traj.direct.df, group==grp, compartment=="rhizosphere")$length,
                                conf.int=TRUE, conf.level=0.95)
  Bac_stage.comp_len.wilcox.df = rbind(Bac_stage.comp_len.wilcox.df, 
                             data.frame(group = grp,
                                        ymax = max(filter(Bac_stage.traj.direct.df, group==grp)$length),
                                        pvalue = comp_len.wilcox$p.value,
                                        W = comp_len.wilcox$statistic,
                                        diff = comp_len.wilcox$estimate))
}

Bac_stage.comp_len.wilcox.df = Bac_stage.comp_len.wilcox.df %>%
  mutate(sig = ifelse(pvalue < 0.001, "p < 0.001",
                      paste("p =", round(pvalue, digits = 3))))

Bac_stage.comp_len.plot = ggplot(data=Bac_stage.traj.direct.df, aes(x=compartment, y=length)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(height=0, width=0.25, size=2) +
  geom_text(data=Bac_stage.comp_len.wilcox.df, aes(label = sig, y=ymax), size=8*5/14, x=1.5) +
  labs(x = "Plant Compartment", y = "Trajectory length") +
  present_theme +
  facet_wrap(~group, nrow=1)

Bac_stage.comp_len.plot

#Collapsing both control and Zeb-treated samples: Length
Bac_coll_stage.comp_len.wilcox.df = data.frame()
Bac_coll_stage_comp_len.wilcox = wilcox.test(
  x=filter(Bac_stage.traj.direct.df, compartment=="rhizoplane")$length,
  y=filter(Bac_stage.traj.direct.df, compartment=="rhizosphere")$length,
  conf.int=TRUE, conf.level=0.95)   
Bac_coll_stage.comp_len.wilcox.df = rbind(Bac_coll_stage.comp_len.wilcox.df,
                                 data.frame(ymax = max(Bac_stage.traj.direct.df$length),
                                            pvalue = Bac_coll_stage_comp_len.wilcox$p.value,
                                            W = Bac_coll_stage_comp_len.wilcox$statistic,
                                            diff = Bac_coll_stage_comp_len.wilcox$estimate))

Bac_coll_stage.comp_len.wilcox.df = Bac_coll_stage.comp_len.wilcox.df %>%   
  mutate(sig = ifelse(pvalue < 0.001, "p < 0.001",                       
                      paste("p =", round(pvalue, digits = 3))))  

Bac_coll_stage.comp_len.plot = 
  ggplot(data=Bac_stage.traj.direct.df, aes(x=compartment, y=length)) +   
  geom_boxplot(outlier.shape = NA, fill = c("#33CCFF","#FF6633")) +   
  geom_jitter(height=0, width=0.25, size=2) +   
  geom_text(data=Bac_coll_stage.comp_len.wilcox.df, aes(label = sig, y=ymax), size=8*5/14, x=1.5) +
  labs(x = "Plant Compartment", y = "Trajectory length") +   
  present_theme

Bac_coll_stage.comp_len.plot
ggsave("Length_Collapsed_Stage_16S.png", Bac_coll_stage.comp_len.plot)
```

```{r}  
#ITS
# Length 
ITS_stage.comp_len.wilcox.df = data.frame() 
for (grp in c("Control", "Zeb")){   
  comp_len.wilcox = wilcox.test(x=filter(ITS_stage.traj.direct.df, treatment==grp, compartment=="rhizoplane")$length,                             
                                y=filter(ITS_stage.traj.direct.df, treatment==grp, compartment=="rhizosphere")$length,                                 
                                conf.int=TRUE, conf.level=0.95)   
  ITS_stage.comp_len.wilcox.df = rbind(ITS_stage.comp_len.wilcox.df,
                                   data.frame(treatment = grp,
                                              ymax = max(filter(ITS_stage.traj.direct.df, treatment==grp)$length),
                                              pvalue = comp_len.wilcox$p.value,
                                              W = comp_len.wilcox$statistic,
                                              diff = comp_len.wilcox$estimate)) 
  }  

ITS_stage.comp_len.wilcox.df = ITS_stage.comp_len.wilcox.df %>%   
  mutate(sig = ifelse(pvalue < 0.001, "p < 0.001",                       
                      paste("p =", round(pvalue, digits = 3))))  

ITS_stage.comp_len.plot = 
  ggplot(data=ITS_stage.traj.direct.df, aes(x=compartment, y=length)) +   
  geom_boxplot(outlier.shape = NA) +   
  geom_jitter(height=0, width=0.25, size=2) +   
  geom_text(data=ITS_stage.comp_len.wilcox.df, aes(label = sig, y=ymax), size=8*5/14, x=1.5) +   
  labs(x = "Plant Compartment", y = "Trajectory length") +   
  present_theme +   
  facet_wrap(~treatment, nrow=1)  

ITS_stage.comp_len.plot

#Collapsing both control and Zeb-treated samples: Length
ITS_coll_stage.comp_len.wilcox.df = data.frame()
ITS_coll_stage_comp_len.wilcox = wilcox.test(
  x=filter(ITS_stage.traj.direct.df, compartment=="rhizoplane")$length,
  y=filter(ITS_stage.traj.direct.df, compartment=="rhizosphere")$length,
  conf.int=TRUE, conf.level=0.95)   
ITS_coll_stage.comp_len.wilcox.df = rbind(ITS_coll_stage.comp_len.wilcox.df,
                                 data.frame(ymax = max(ITS_stage.traj.direct.df$length),
                                            pvalue = ITS_coll_stage_comp_len.wilcox$p.value,
                                            W = ITS_coll_stage_comp_len.wilcox$statistic,
                                            diff = ITS_coll_stage_comp_len.wilcox$estimate))

ITS_coll_stage.comp_len.wilcox.df = ITS_coll_stage.comp_len.wilcox.df %>%   
  mutate(sig = ifelse(pvalue < 0.001, "p < 0.001",                       
                      paste("p =", round(pvalue, digits = 3))))  

ITS_coll_stage.comp_len.plot = 
  ggplot(data=ITS_stage.traj.direct.df, aes(x=compartment, y=length)) +   
  geom_boxplot(outlier.shape = NA, fill = c("#33CCFF","#FF6633")) +   
  geom_jitter(height=0, width=0.25, size=2) +   
  geom_text(data=ITS_coll_stage.comp_len.wilcox.df, aes(label = sig, y=ymax), size=8*5/14, x=1.5) +
  labs(x = "Plant Compartment", y = "Trajectory length") +   
  present_theme

ITS_coll_stage.comp_len.plot
ggsave("Length_Collapsed_Stage_ITS.png", ITS_coll_stage.comp_len.plot)
```
Rhizoplane has significantly more trajectory length (but not directionality) than rhizosphere regardless of the treatment.

# Super Figure
```{r}
ITS_row <- cowplot::plot_grid(ITS_coll_time.comp_dir.plot +
                         lims(y = c(0.305, 0.375)) +
                         ggtitle("Harvest at same timepoint"),
                       ITS_coll_stage.comp_dir.plot +
                         lims(y = c(0.305, 0.375)) +
                         ggtitle("Harvest at same growth stage"),
                       ncol = 2)

BAC_row <- cowplot::plot_grid(Bac_coll_time.comp_dir.plot +
                            lims(y = c(0.30, 0.44)) +
                            ggtitle("Harvest at same timepoint"),
                          Bac_coll_stage.comp_dir.plot +
                            lims(y = c(0.30, 0.44)) +
                            ggtitle("Harvest at same growth stage"),
                          ncol = 2)

A_title <- ggdraw() +
  draw_label("A. Fungi", fontface = "bold", x = 0.02, hjust = 0)

B_title <- ggdraw() +
  draw_label("B. Bacteria", fontface = "bold", x = 0.02, hjust = 0)

final_plot <- plot_grid(
  A_title,
  ITS_row,
  B_title,
  BAC_row,
  ncol = 1,
  rel_heights = c(0.05, 1, 0.05, 1))
final_plot
ggsave("Full_Trajectory_Collapsed.png", width = 8, height = 8, dpi = 300)
```
# Community dissimilarity over time and life stage

Now lets see how different the Zeb treated and untreated communities are over time. Are they becoming less similar over time or is there a peak in dissimilarity?

#### ITS
## Across timepoints

First, lets look at dissimilarity between treated and untreated plants within each timepoint. Note that life stage will likely be different at each timepoint.

```{r}         
# Get new metadata for each sample pair 
ITS_sample_1.meta = data.frame(sample_data(ITS_time.rare.physeq)) %>% 
  select(SampleID, treatment, age_days, compartment) %>%
  dplyr::rename(Sample_1 = SampleID,
         group_1 = treatment,          
         stage_1 = age_days,          
         compartment_1 = compartment)   

ITS_sample_2.meta = data.frame(sample_data(ITS_time.rare.physeq)) %>%   
  select(SampleID, treatment, age_days, compartment) %>%   
  dplyr::rename(Sample_2 = SampleID,          
         group_2 = treatment,          
         stage_2 = age_days,          
         compartment_2 = compartment)  

# Get Bray-Curtis dissimilarity between each plant in dataframe format and add in metadata 

ITS_time.BC.dist.mat = as.matrix(ITS_time.BC.dist) 
ITS_time.BC.dist.mat[upper.tri(ITS_time.BC.dist.mat, diag = TRUE)] = NA 
ITS_time.BC.dist.df = data.frame(ITS_time.BC.dist.mat) %>%   
  tibble::rownames_to_column(var="Sample_1") %>%   
  tidyr::gather(key="Sample_2", value="BC", -Sample_1) %>%   
  filter(!is.na(BC)) %>%   
  left_join(ITS_sample_1.meta) %>%   
  left_join(ITS_sample_2.meta)  

# Filter to same timepoint, compartment but different treatment. 
ITS_sub.time.BC.dist.df = ITS_time.BC.dist.df %>%   
  filter(stage_1 == stage_2,
         compartment_1 == compartment_2,          
         group_1 != group_2) %>%   
  mutate(day = factor(gsub("day", " ", stage_1), 
                      levels=c("3", "7", "14", "21", "35", "49", "63"))) 

# Run Kruskal-Wallis tests and Dunn's Tests 
ITS_time.BC.dist.dunn.cld.df = data.frame() 
for (com in c("rhizoplane", "rhizosphere")){
  ITS_time.BC.dist.kruskal = kruskal.test(BC ~ day, data=filter(ITS_sub.time.BC.dist.df, compartment_1==com))   
  print(com)   
  print(ITS_time.BC.dist.kruskal)   
  ITS_time.BC.dist.dunn = FSA::dunnTest(BC ~ day, data=filter(ITS_sub.time.BC.dist.df, compartment_1==com), method="bh")   
  ITS_time.BC.dist.dunn.cld = rcompanion::cldList(comparison = ITS_time.BC.dist.dunn$res$Comparison,                                               
                                              p.value = ITS_time.BC.dist.dunn$res$P.adj,
                                              threshold = 0.05) %>%     
    mutate(day = Group,            
           compartment_1 = com)   
  ITS_time.BC.dist.dunn.cld.df = rbind(ITS_time.BC.dist.dunn.cld.df, ITS_time.BC.dist.dunn.cld) 
  }
```
```{r}         
# Plot 
ITS_time.BC.dist.plot = ggplot(data=ITS_sub.time.BC.dist.df, aes(x=day, y=BC)) +   
  geom_boxplot(outlier.shape = NA) +   
  geom_jitter(height=0, width = 0.25) +   
  geom_text(data=ITS_time.BC.dist.dunn.cld.df, aes(x=day, label=Letter), y = 1) +   
  lims(y=c(0,1)) +   
  present_theme +   
  facet_wrap(~compartment_1)  

ITS_time.BC.dist.plot
ggsave("ITS_Dissimilarity_Time.png")
```
## Across life stage

Next, lets look across the life stages.

```{r}         
# Get new metadata for each sample pair 
ITS_sample_1.meta = data.frame(sample_data(ITS_stage.rare.physeq)) %>%   
  select(SampleID, treatment, Development, compartment) %>%   
  dplyr::rename(Sample_1 = SampleID,          
         group_1 = treatment,          
         stage_1 = Development,          
         compartment_1 = compartment)   

ITS_sample_2.meta = data.frame(sample_data(ITS_stage.rare.physeq)) %>%   
  select(SampleID, treatment, Development, compartment) %>%   
  dplyr::rename(Sample_2 = SampleID,          
         group_2 = treatment,          
         stage_2 = Development,          
         compartment_2 = compartment)  

# Get Bray-Curtis dissimilarity between each plant in dataframe format and add in metadata 
ITS_stage.BC.dist.mat = as.matrix(ITS_stage.BC.dist) 
ITS_stage.BC.dist.mat[upper.tri(ITS_stage.BC.dist.mat, diag = TRUE)] = NA 
ITS_stage.BC.dist.df = data.frame(ITS_stage.BC.dist.mat) %>%   
  tibble::rownames_to_column(var="Sample_1") %>%   
  tidyr::gather(key="Sample_2", value="BC", -Sample_1) %>%   
  filter(!is.na(BC)) %>%   
  left_join(ITS_sample_1.meta) %>%   
  left_join(ITS_sample_2.meta) 

# Filter to same stagepoint, compartment but different treatment. 
ITS_sub.stage.BC.dist.df = ITS_stage.BC.dist.df %>%   
  filter(stage_1 == stage_2,          
         compartment_1 == compartment_2,          
         group_1 != group_2) %>%   
  mutate(stage = factor(stage_1, 
                        levels=c("V1", "V2", "V3", "V4", "R1", "R4", "R7")))  

# Run Kruskal-Wallis tests and Dunn's Tests 
ITS_stage.BC.dist.dunn.cld.df = data.frame() 
for (com in c("rhizoplane", "rhizosphere")){   
  ITS_stage.BC.dist.kruskal = kruskal.test(BC ~ stage, data=filter(ITS_sub.stage.BC.dist.df, compartment_1==com))   
  print(com)   
  print(ITS_stage.BC.dist.kruskal)   
  ITS_stage.BC.dist.dunn = FSA::dunnTest(BC ~ stage, data=filter(ITS_sub.stage.BC.dist.df, compartment_1==com), method="bh")   
  ITS_stage.BC.dist.dunn.cld = rcompanion::cldList(comparison = ITS_stage.BC.dist.dunn$res$Comparison,                                               
                                               p.value = ITS_stage.BC.dist.dunn$res$P.adj,
                                               threshold = 0.05) %>%     
    mutate(stage = Group,            
           compartment_1 = com)   
  ITS_stage.BC.dist.dunn.cld.df = rbind(ITS_stage.BC.dist.dunn.cld.df, ITS_stage.BC.dist.dunn.cld)
  }
```


```{r}         
# Plot 
ITS_stage.BC.dist.plot = ggplot(data=ITS_sub.stage.BC.dist.df, aes(x=stage, y=BC)) +   
  geom_boxplot(outlier.shape = NA) +   
  geom_jitter(height=0, width = 0.25) +   
  geom_text(data=ITS_stage.BC.dist.dunn.cld.df, aes(x=stage, label=Letter), y = 1) +   
  lims(y=c(0,1)) +   
  present_theme +   
  facet_wrap(~compartment_1)  

ITS_stage.BC.dist.plot
ggsave("ITS_Dissimilarity_Stage.png")
```
#### 16S
## Across timepoints

First, lets look at dissimilarity between treated and untreated plants within each timepoint. Note that life stage will likely be different at each timepoint.

```{r}         
Bac_sample_1.meta = data.frame(sample_data(Bac_time.rare.physeq)) %>%
  select(SampleID, group, stage, compartment) %>%
  rename(Sample_1 = SampleID,
         group_1 = group,
         stage_1 = stage,
         compartment_1 = compartment)


Bac_sample_2.meta = data.frame(sample_data(Bac_time.rare.physeq)) %>%
  select(SampleID, group, stage, compartment) %>%
  rename(Sample_2 = SampleID,
         group_2 = group,
         stage_2 = stage,
         compartment_2 = compartment)

# Get Bray-Curtis dissimilarity between each plant in dataframe format and add in metadata
Bac_time.BC.dist.mat = as.matrix(Bac_time.BC.dist)
Bac_time.BC.dist.mat[upper.tri(Bac_time.BC.dist.mat, diag = TRUE)] = NA
Bac_time.BC.dist.df = data.frame(Bac_time.BC.dist.mat) %>%
  tibble::rownames_to_column(var="Sample_1") %>%
  tidyr::gather(key="Sample_2", value="BC", -Sample_1) %>%
  filter(!is.na(BC)) %>%
  left_join(Bac_sample_1.meta) %>%
  left_join(Bac_sample_2.meta)

# Filter to same timepoint, compartment but different treatment.
Bac_sub.time.BC.dist.df = Bac_time.BC.dist.df %>%
  filter(stage_1 == stage_2,
         compartment_1 == compartment_2,
         group_1 != group_2) %>%
  mutate(day = factor(gsub("day", "", stage_1), levels=c("3", "7", "14", "21", "35", "49", "63")))

# Run Kruskal-Wallis tests and Dunn's Tests
Bac_time.BC.dist.dunn.cld.df = data.frame()
for (com in c("rhizoplane", "rhizosphere")){
  Bac_time.BC.dist.kruskal = kruskal.test(BC ~ day, data=filter(Bac_sub.time.BC.dist.df, compartment_1==com))
  print(com)
  print(Bac_time.BC.dist.kruskal)
  Bac_time.BC.dist.dunn = FSA::dunnTest(BC ~ day, data=filter(Bac_sub.time.BC.dist.df, compartment_1==com), method="bh")
  Bac_time.BC.dist.dunn.cld = rcompanion::cldList(comparison = Bac_time.BC.dist.dunn$res$Comparison,
                                              p.value = Bac_time.BC.dist.dunn$res$P.adj,
                                              threshold = 0.05) %>%
    mutate(day = Group,
           compartment_1 = com)
  Bac_time.BC.dist.dunn.cld.df = rbind(Bac_time.BC.dist.dunn.cld.df, Bac_time.BC.dist.dunn.cld)
}
```
```{r}
Bac_time.BC.dist.plot = ggplot(data=Bac_sub.time.BC.dist.df, aes(x=day, y=BC)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(height=0, width = 0.25) +
  geom_text(data=Bac_time.BC.dist.dunn.cld.df, aes(x=day, label=Letter), y = 1) +
  lims(y=c(0,1)) +
  present_theme +
  facet_wrap(~compartment_1)

Bac_time.BC.dist.plot
ggsave("16S_Dissimilarity_Time.png")
```
##Across life stage
Next, let's look across the life stages

```{r}
# Get new metadata for each sample pair
Bac_sample_1.meta = data.frame(sample_data(Bac_stage.rare.physeq)) %>%
  select(SampleID, group, stage, compartment) %>%
  rename(Sample_1 = SampleID,
         group_1 = group,
         stage_1 = stage,
         compartment_1 = compartment)


Bac_sample_2.meta = data.frame(sample_data(Bac_stage.rare.physeq)) %>%
  select(SampleID, group, stage, compartment) %>%
  rename(Sample_2 = SampleID,
         group_2 = group,
         stage_2 = stage,
         compartment_2 = compartment)

# Get Bray-Curtis dissimilarity between each plant in dataframe format and add in metadata
Bac_stage.BC.dist.mat = as.matrix(Bac_stage.BC.dist)
Bac_stage.BC.dist.mat[upper.tri(Bac_stage.BC.dist.mat, diag = TRUE)] = NA
Bac_stage.BC.dist.df = data.frame(Bac_stage.BC.dist.mat) %>%
  tibble::rownames_to_column(var="Sample_1") %>%
  tidyr::gather(key="Sample_2", value="BC", -Sample_1) %>%
  filter(!is.na(BC)) %>%
  left_join(Bac_sample_1.meta) %>%
  left_join(Bac_sample_2.meta)

# Filter to same stagepoint, compartment but different treatment.
Bac_sub.stage.BC.dist.df = Bac_stage.BC.dist.df %>%
  filter(stage_1 == stage_2,
         compartment_1 == compartment_2,
         group_1 != group_2) %>%
  mutate(stage = factor(stage_1, levels=c("V1", "V2", "V3", "V4", "R1", "R4", "R7")))

# Run Kruskal-Wallis tests and Dunn's Tests
Bac_stage.BC.dist.dunn.cld.df = data.frame()
for (com in c("rhizoplane", "rhizosphere")){
  Bac_stage.BC.dist.kruskal = kruskal.test(BC ~ stage, data=filter(Bac_sub.stage.BC.dist.df, compartment_1==com))
  print(com)
  print(Bac_stage.BC.dist.kruskal)
  Bac_stage.BC.dist.dunn = FSA::dunnTest(BC ~ stage, data=filter(Bac_sub.stage.BC.dist.df, compartment_1==com), method="bh")
  Bac_stage.BC.dist.dunn.cld = rcompanion::cldList(comparison = Bac_stage.BC.dist.dunn$res$Comparison,
                                              p.value = Bac_stage.BC.dist.dunn$res$P.adj,
                                              threshold = 0.05) %>%
    mutate(stage = Group,
           compartment_1 = com)
  Bac_stage.BC.dist.dunn.cld.df = rbind(Bac_stage.BC.dist.dunn.cld.df, Bac_stage.BC.dist.dunn.cld)
}
```
```{r}
# Plot
Bac_stage.BC.dist.plot = ggplot(data=Bac_sub.stage.BC.dist.df, aes(x=stage, y=BC)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(height=0, width = 0.25) +
  geom_text(data=Bac_stage.BC.dist.dunn.cld.df, aes(x=stage, label=Letter), y = 1) +
  lims(y=c(0,1)) +
  present_theme +
  facet_wrap(~compartment_1)

Bac_stage.BC.dist.plot
ggsave("16S_Dissimilarity_Stage.png")
```
## Final Figures

```{r}         
#TIME
  #ITS RHIZOPLANE
rp_ITS_sub.time.BC.dist.df <- ITS_sub.time.BC.dist.df %>% 
  filter(compartment_1 == "rhizoplane")
rp_ITS_sub.time.BC.dist.df

rp_ITS_time.BC.dist.dunn.cld.df <- ITS_time.BC.dist.dunn.cld.df %>% 
  filter(compartment_1 == "rhizoplane")
rp_ITS_time.BC.dist.dunn.cld.df

rp_ITS_sub.time.BC.dist.df$time.series <- "A2. ITS"

rp_ITS_time.BC.dist.plot = ggplot(data=rp_ITS_sub.time.BC.dist.df, aes(x=day, y=BC)) +   
  geom_boxplot(outlier.shape = NA, fill = "#FCE8FC") +   
  geom_jitter(height=0, width = 0.25) +   
  geom_text(data=rp_ITS_time.BC.dist.dunn.cld.df, aes(x=day, label=Letter), y = 1) +   
  lims(y=c(0,1)) +   
  present_theme +
  facet_grid(.~time.series)
rp_ITS_time.BC.dist.plot

  #ITS RHIZOSPHERE
rs_ITS_sub.time.BC.dist.df <- ITS_sub.time.BC.dist.df %>% 
  filter(compartment_1 == "rhizosphere")
rs_ITS_sub.time.BC.dist.df

rs_ITS_time.BC.dist.dunn.cld.df <- ITS_time.BC.dist.dunn.cld.df %>% 
  filter(compartment_1 == "rhizosphere")
rs_ITS_time.BC.dist.dunn.cld.df

rs_ITS_sub.time.BC.dist.df$time.series <- "A2. ITS"

rs_ITS_time.BC.dist.plot = ggplot(data=rs_ITS_sub.time.BC.dist.df, aes(x=day, y=BC)) +   
  geom_boxplot(outlier.shape = NA, fill = "#FCE8FC") +   
  geom_jitter(height=0, width = 0.25) +   
  geom_text(data=rs_ITS_time.BC.dist.dunn.cld.df, aes(x=day, label=Letter), y = 1) +   
  lims(y=c(0,1)) +   
  present_theme +
  facet_grid(.~time.series)
rs_ITS_time.BC.dist.plot

  #16S RHIZOPLAN
rp_Bac_sub.time.BC.dist.df <- Bac_sub.time.BC.dist.df %>% 
  filter(compartment_1 == "rhizoplane")
rp_Bac_sub.time.BC.dist.df

rp_Bac_time.BC.dist.dunn.cld.df <- Bac_time.BC.dist.dunn.cld.df %>% 
  filter(compartment_1 == "rhizoplane")
rp_Bac_time.BC.dist.dunn.cld.df

rp_Bac_sub.time.BC.dist.df$time.series <- "A1. 16S"

rp_Bac_time.BC.dist.plot = ggplot(data=rp_Bac_sub.time.BC.dist.df, aes(x=day, y=BC)) +   
  geom_boxplot(outlier.shape = NA, fill = "#E8FCE9") +   
  geom_jitter(height=0, width = 0.25) +   
  geom_text(data=rp_Bac_time.BC.dist.dunn.cld.df, aes(x=day, label=Letter), y = 1) +   
  lims(y=c(0,1)) +   
  present_theme +
  facet_grid(.~time.series)
rp_Bac_time.BC.dist.plot

  #16S RHIZOSPHERE
rs_Bac_sub.time.BC.dist.df <- Bac_sub.time.BC.dist.df %>% 
  filter(compartment_1 == "rhizosphere")
rs_Bac_sub.time.BC.dist.df

rs_Bac_time.BC.dist.dunn.cld.df <- Bac_time.BC.dist.dunn.cld.df %>% 
  filter(compartment_1 == "rhizosphere")
rs_Bac_time.BC.dist.dunn.cld.df

rs_Bac_sub.time.BC.dist.df$time.series <- "A1. 16S"

rs_Bac_time.BC.dist.plot = ggplot(data=rs_Bac_sub.time.BC.dist.df, aes(x=day, y=BC)) +   
  geom_boxplot(outlier.shape = NA, fill= "#E8FCE9") +   
  geom_jitter(height=0, width = 0.25) +   
  geom_text(data=rs_ITS_time.BC.dist.dunn.cld.df, aes(x=day, label=Letter), y = 1) +   
  lims(y=c(0,1)) +   
  present_theme +
  facet_grid(.~ time.series)
rs_Bac_time.BC.dist.plot

#STAGE
  #ITS RHIZOPLANE
rp_ITS_sub.stage.BC.dist.df <- ITS_sub.stage.BC.dist.df %>% 
  filter(compartment_1 == "rhizoplane")
rp_ITS_sub.stage.BC.dist.df

rp_ITS_stage.BC.dist.dunn.cld.df <- ITS_stage.BC.dist.dunn.cld.df %>% 
  filter(compartment_1 == "rhizoplane")
rp_ITS_stage.BC.dist.dunn.cld.df

rp_ITS_sub.stage.BC.dist.df$growth.series <- "B2. ITS"

rp_ITS_stage.BC.dist.plot = ggplot(data=rp_ITS_sub.stage.BC.dist.df, aes(x=stage, y=BC)) +   geom_boxplot(outlier.shape = NA, fill = "#F297F1") +   
  geom_jitter(height=0, width = 0.25) +   
  geom_text(data=rp_ITS_stage.BC.dist.dunn.cld.df, aes(x=stage, label=Letter), y = 1) +   
  lims(y=c(0,1)) +   
  present_theme +
  facet_grid(.~ growth.series)
rp_ITS_stage.BC.dist.plot

  #ITS RHIZOSPHERE
rs_ITS_sub.stage.BC.dist.df <- ITS_sub.stage.BC.dist.df %>% 
  filter(compartment_1 == "rhizosphere")
rs_ITS_sub.stage.BC.dist.df

rs_ITS_stage.BC.dist.dunn.cld.df <- ITS_stage.BC.dist.dunn.cld.df %>% 
  filter(compartment_1 == "rhizosphere")
rs_ITS_stage.BC.dist.dunn.cld.df

rs_ITS_sub.stage.BC.dist.df$growth.series <- "B2. ITS"

rs_ITS_stage.BC.dist.plot = ggplot(data=rs_ITS_sub.stage.BC.dist.df, aes(x=stage, y=BC)) +   geom_boxplot(outlier.shape = NA, fill = "#F297F1") +   
  geom_jitter(height=0, width = 0.25) +   
  geom_text(data=rs_ITS_stage.BC.dist.dunn.cld.df, aes(x=stage, label=Letter), y = 1) +   
  lims(y=c(0,1)) +   
  present_theme + 
  facet_grid(. ~ growth.series)
rs_ITS_stage.BC.dist.plot

  #16S RHIZOPLAN
rp_Bac_sub.stage.BC.dist.df <- Bac_sub.stage.BC.dist.df %>% 
  filter(compartment_1 == "rhizoplane")
rp_Bac_sub.stage.BC.dist.df

rp_Bac_stage.BC.dist.dunn.cld.df <- Bac_stage.BC.dist.dunn.cld.df %>% 
  filter(compartment_1 == "rhizoplane")
rp_Bac_stage.BC.dist.dunn.cld.df

rp_Bac_sub.stage.BC.dist.df$growth.series <- "B1. 16S"

rp_Bac_stage.BC.dist.plot = ggplot(data=rp_Bac_sub.stage.BC.dist.df, aes(x=stage, y=BC)) +   geom_boxplot(outlier.shape = NA, fill = "#97F298") +   
  geom_jitter(height=0, width = 0.25) +   
  geom_text(data=rp_Bac_stage.BC.dist.dunn.cld.df, aes(x=stage, label=Letter), y = 1) +   
  lims(y=c(0,1)) +   
  present_theme +
  facet_grid(. ~ growth.series)
rp_Bac_stage.BC.dist.plot

  #16S RHIZOSPHERE
rs_Bac_sub.stage.BC.dist.df <- Bac_sub.stage.BC.dist.df %>% 
  filter(compartment_1 == "rhizosphere")
rs_Bac_sub.stage.BC.dist.df

rs_Bac_stage.BC.dist.dunn.cld.df <- Bac_stage.BC.dist.dunn.cld.df %>% 
  filter(compartment_1 == "rhizosphere")
rs_Bac_stage.BC.dist.dunn.cld.df

rs_Bac_sub.stage.BC.dist.df$growth.series <- "B1. 16S"

rs_Bac_stage.BC.dist.plot = ggplot(data=rs_Bac_sub.stage.BC.dist.df, aes(x=stage, y=BC)) +   geom_boxplot(outlier.shape = NA, fill = "#97F298") +   
  geom_jitter(height=0, width = 0.25) +   
  geom_text(data=rs_Bac_stage.BC.dist.dunn.cld.df, aes(x=stage, label=Letter), y = 1) +   
  lims(y=c(0,1)) +   
  present_theme +
  facet_grid(. ~ growth.series)
  
rs_Bac_stage.BC.dist.plot

  #Assemble

A_title <- ggdraw() +
  draw_label("A. Harvest at same timepoint", fontface = "bold", x = 0.02, 
             hjust = 0, size = 10)

B_title <- ggdraw() +
  draw_label("B. Harvest at same growth stage", fontface = "bold", x = 0.02, 
             hjust = 0, size = 10)

RS_Time_row <- cowplot::plot_grid(rs_Bac_time.BC.dist.plot,
                                  rs_ITS_time.BC.dist.plot,
                                  ncol = 2)
RS_Growth_row <- cowplot::plot_grid(rs_Bac_stage.BC.dist.plot,
                                    rs_ITS_stage.BC.dist.plot,
                                    ncol = 2)
RS_Dissimilarity_plot <- plot_grid(A_title,
                                   RS_Time_row,
                                   B_title,
                                   RS_Growth_row,
                                   ncol = 1,
                                   rel_heights = c(0.05, 1, 0.05, 1))
RS_Dissimilarity_plot
ggsave("RHIZOSPHERE_Dissimilarity_full.png")

RP_Time_row <- cowplot::plot_grid(rp_Bac_time.BC.dist.plot,
                                  rp_ITS_time.BC.dist.plot,
                                  ncol = 2)
RP_Growth_row <- cowplot::plot_grid(rp_Bac_stage.BC.dist.plot,
                                    rp_ITS_stage.BC.dist.plot,
                                    ncol = 2)
RP_Dissimilarity_plot <- plot_grid(A_title,
                                   RP_Time_row,
                                   B_title,
                                   RP_Growth_row,
                                   ncol = 1,
                                   rel_heights = c(0.05, 1, 0.05, 1))
RP_Dissimilarity_plot
ggsave("RHIZOPLANE_Dissimilarity_full.png")

```

##Tables of metrics: cumulative distance, directionality, betadispersion
```{r} 
#TIME
  #ITS
nperm <- 100
res_list <- list()

for (i in seq_len(nperm)) {
  set.seed(i + 42)

  Iter.meta <- data.frame(sample_data(ITS_time.rare.physeq)) %>%
    select(SampleID, treatment, series, age_days, plant_rep, compartment) %>%
    mutate(Stage_num = as.numeric(age_days)) %>%
    group_by(treatment, series, age_days, compartment) %>%
    sample_n(1) %>%
    ungroup() %>%
    mutate(PlantID = paste(treatment, compartment, series, "Iteration", i, sep = "_")) %>%
    arrange(PlantID, Stage_num)

  D <- as.matrix(ITS_time.BC.dist)
  D <- as.dist(D[Iter.meta$SampleID, Iter.meta$SampleID])

  #Directionality and cumulative distance
  traj <- defineTrajectories(D, Iter.meta$PlantID, Iter.meta$Stage_num)
  met <- trajectoryMetrics(traj) %>%
    dplyr::rename(PlantID = trajectory) %>%
    select(PlantID, n_timepoints = n, cumulative_distance = length, directionality)

  #Beta dispersion
  grp <- interaction(Iter.meta$treatment, Iter.meta$compartment, drop = TRUE)
  bd <- betadisper(D, grp, type = "centroid")

  disp_plant <- data.frame(PlantID = Iter.meta$PlantID,
                           dist_to_centroid = bd$distances) %>%
    group_by(PlantID) %>%
    summarise(beta_disp = mean(dist_to_centroid), .groups = "drop")

  #Replicate number
  n_grp <- Iter.meta %>%
    group_by(treatment, compartment) %>%
    summarise(n_trajectories = n_distinct(PlantID),
              n_samples = n(),
              .groups = "drop")

  #Build the table
  res_list[[i]] <- met %>%
    left_join(disp_plant, by = "PlantID") %>%
    left_join(Iter.meta %>% select(PlantID, treatment, series, compartment) %>% unique(),
              by = "PlantID") %>%
    left_join(n_grp, by = c("treatment", "compartment")) %>%
    mutate(Iteration = i)
}

ITS_time.traj.all.df <- bind_rows(res_list)

#Summary table
ITS_time.traj.summary <- ITS_time.traj.all.df %>%
  group_by(treatment, compartment, Iteration) %>%
  summarise(across(c(directionality, cumulative_distance, beta_disp), mean),
            n_trajectories = first(n_trajectories),
            n_samples = first(n_samples),
            .groups = "drop") %>%
  group_by(treatment, compartment) %>%
  summarise(across(c(directionality, cumulative_distance, beta_disp),
                   list(mean = mean, sd = sd,
                        q025 = ~quantile(.x, 0.025), q975 = ~quantile(.x, 0.975)),
                   .names = "{.col}_{.fn}"),
            n_trajectories = mean(n_trajectories),
            n_samples = mean(n_samples),
            .groups = "drop")

ITS_time.traj.summary

## Export
write.csv(ITS_time.traj.all.df, "ITS_time_traj_all_iterations.csv", row.names = FALSE)
write.csv(ITS_time.traj.summary, "ITS_time_traj_summary.csv", row.names = FALSE)
 
  #16S
nperm <- 100
res_list <- list()

for (i in seq_len(nperm)) {
  set.seed(i + 42)

  Iter.meta <- data.frame(sample_data(Bac_time.rare.physeq)) %>%
    select(SampleID, group, series, stage, plant_rep, compartment) %>%
    mutate(Stage_num = as.numeric(gsub("day", "", stage))) %>%
    group_by(group, series, stage, compartment) %>%
    sample_n(1) %>%
    ungroup() %>%
    mutate(PlantID = paste(group, compartment, series, "Iteration", i, sep = "_")) %>%
    arrange(PlantID, Stage_num)

  D <- as.matrix(Bac_time.BC.dist)
  D <- as.dist(D[Iter.meta$SampleID, Iter.meta$SampleID])

  #Directionality and cumulative distance
  traj <- defineTrajectories(D, Iter.meta$PlantID, Iter.meta$Stage_num)
  met <- trajectoryMetrics(traj) %>%
    dplyr::rename(PlantID = trajectory) %>%
    select(PlantID, n_timepoints = n, cumulative_distance = length, directionality)

  #Beta dispersion
  grp <- interaction(Iter.meta$group, Iter.meta$compartment, drop = TRUE)
  bd <- betadisper(D, grp, type = "centroid")

  disp_plant <- data.frame(PlantID = Iter.meta$PlantID,
                           dist_to_centroid = bd$distances) %>%
    group_by(PlantID) %>%
    summarise(beta_disp = mean(dist_to_centroid), .groups = "drop")

  #Replicate number
  n_grp <- Iter.meta %>%
    group_by(group, compartment) %>%
    summarise(n_trajectories = n_distinct(PlantID),
              n_samples = n(),
              .groups = "drop")

  #Build the table
  res_list[[i]] <- met %>%
    left_join(disp_plant, by = "PlantID") %>%
    left_join(Iter.meta %>% select(PlantID, group, series, compartment) %>% unique(),
              by = "PlantID") %>%
    left_join(n_grp, by = c("group", "compartment")) %>%
    mutate(Iteration = i)
}

Bac_time.traj.all.df <- bind_rows(res_list)

#Summary table
Bac_time.traj.summary <- Bac_time.traj.all.df %>%
  group_by(group, compartment, Iteration) %>%
  summarise(across(c(directionality, cumulative_distance, beta_disp), mean),
            n_trajectories = first(n_trajectories),
            n_samples = first(n_samples),
            .groups = "drop") %>%
  group_by(group, compartment) %>%
  summarise(across(c(directionality, cumulative_distance, beta_disp),
                   list(mean = mean, sd = sd,
                        q025 = ~quantile(.x, 0.025), q975 = ~quantile(.x, 0.975)),
                   .names = "{.col}_{.fn}"),
            n_trajectories = mean(n_trajectories),
            n_samples = mean(n_samples),
            .groups = "drop")

Bac_time.traj.summary

## Export
write.csv(Bac_time.traj.all.df, "Bac_time_traj_all_iterations.csv", row.names = FALSE)
write.csv(Bac_time.traj.summary, "Bac_time_traj_summary.csv", row.names = FALSE)

#STAGE
  #ITS
stage_num.df = data.frame(Stage_num = c(1,2,3,4,5,6,7),                           
                          Development = c("V1", "V2", "V3", "V4", "R1", "R4", "R7"))  
nperm <- 100
res_list <- list()

for (i in seq_len(nperm)) {
  set.seed(i + 42)

  Iter.meta <- data.frame(sample_data(ITS_stage.rare.physeq)) %>%
    select(SampleID, treatment, series, Development, plant_rep, compartment) %>%
    left_join(stage_num.df) %>%
    group_by(treatment, series, Development, compartment) %>%
    sample_n(1) %>%
    ungroup() %>%
    mutate(PlantID = paste(treatment, compartment, series, "Iteration", i, sep = "_")) %>%
    arrange(PlantID, Stage_num)

  D <- as.matrix(ITS_stage.BC.dist)
  D <- as.dist(D[Iter.meta$SampleID, Iter.meta$SampleID])

  #Directionality and cumulative distance
  traj <- defineTrajectories(D, Iter.meta$PlantID, Iter.meta$Stage_num)
  met <- trajectoryMetrics(traj) %>%
    dplyr::rename(PlantID = trajectory) %>%
    select(PlantID, n_stagepoints = n, cumulative_distance = length, directionality)

  #Beta dispersion
  grp <- interaction(Iter.meta$treatment, Iter.meta$compartment, drop = TRUE)
  bd <- betadisper(D, grp, type = "centroid")

  disp_plant <- data.frame(PlantID = Iter.meta$PlantID,
                           dist_to_centroid = bd$distances) %>%
    group_by(PlantID) %>%
    summarise(beta_disp = mean(dist_to_centroid), .groups = "drop")

  #Replicate number
  n_grp <- Iter.meta %>%
    group_by(treatment, compartment) %>%
    summarise(n_trajectories = n_distinct(PlantID),
              n_samples = n(),
              .groups = "drop")

  #Build the table
  res_list[[i]] <- met %>%
    left_join(disp_plant, by = "PlantID") %>%
    left_join(Iter.meta %>% select(PlantID, treatment, series, compartment) %>% unique(),
              by = "PlantID") %>%
    left_join(n_grp, by = c("treatment", "compartment")) %>%
    mutate(Iteration = i)
}

ITS_stage.traj.all.df <- bind_rows(res_list)

#Summary table
ITS_stage.traj.summary <- ITS_stage.traj.all.df %>%
  group_by(treatment, compartment, Iteration) %>%
  summarise(across(c(directionality, cumulative_distance, beta_disp), mean),
            n_trajectories = first(n_trajectories),
            n_samples = first(n_samples),
            .groups = "drop") %>%
  group_by(treatment, compartment) %>%
  summarise(across(c(directionality, cumulative_distance, beta_disp),
                   list(mean = mean, sd = sd,
                        q025 = ~quantile(.x, 0.025), q975 = ~quantile(.x, 0.975)),
                   .names = "{.col}_{.fn}"),
            n_trajectories = mean(n_trajectories),
            n_samples = mean(n_samples),
            .groups = "drop")

ITS_stage.traj.summary

## Export
write.csv(ITS_stage.traj.all.df, "ITS_stage_traj_all_iterations.csv", row.names = FALSE)
write.csv(ITS_stage.traj.summary, "ITS_stage_traj_summary.csv", row.names = FALSE)
  
  #16S

stage_num.df = data.frame(Stage_num = c(1,2,3,4,5,6,7),
                          stage = c("V1", "V2", "V3", "V4", "R1", "R4", "R7"))
nperm <- 100
res_list <- list()

for (i in seq_len(nperm)) {
  set.seed(i + 42)

  Iter.meta <- data.frame(sample_data(Bac_stage.rare.physeq)) %>%
    select(SampleID, group, series, stage, plant_rep, compartment) %>%
    left_join(stage_num.df) %>%
    group_by(group, series, stage, compartment) %>%
    sample_n(1) %>%
    ungroup() %>%
    mutate(PlantID = paste(group, compartment, series, "Iteration", i, sep = "_")) %>%
    arrange(PlantID, Stage_num)

  D <- as.matrix(Bac_stage.BC.dist)
  D <- as.dist(D[Iter.meta$SampleID, Iter.meta$SampleID])

  #Directionality and cumulative distance
  traj <- defineTrajectories(D, Iter.meta$PlantID, Iter.meta$Stage_num)
  met <- trajectoryMetrics(traj) %>%
    dplyr::rename(PlantID = trajectory) %>%
    select(PlantID, n_timepoints = n, cumulative_distance = length, directionality)

  #Beta dispersion
  grp <- interaction(Iter.meta$group, Iter.meta$compartment, drop = TRUE)
  bd <- betadisper(D, grp, type = "centroid")

  disp_plant <- data.frame(PlantID = Iter.meta$PlantID,
                           dist_to_centroid = bd$distances) %>%
    group_by(PlantID) %>%
    summarise(beta_disp = mean(dist_to_centroid), .groups = "drop")

  #Replicate number
  n_grp <- Iter.meta %>%
    group_by(group, compartment) %>%
    summarise(n_trajectories = n_distinct(PlantID),
              n_samples = n(),
              .groups = "drop")

  #Build the table
  res_list[[i]] <- met %>%
    left_join(disp_plant, by = "PlantID") %>%
    left_join(Iter.meta %>% select(PlantID, group, series, compartment) %>% unique(),
              by = "PlantID") %>%
    left_join(n_grp, by = c("group", "compartment")) %>%
    mutate(Iteration = i)
}

Bac_stage.traj.all.df <- bind_rows(res_list)

#Summary table
Bac_stage.traj.summary <- Bac_stage.traj.all.df %>%
  group_by(group, compartment, Iteration) %>%
  summarise(across(c(directionality, cumulative_distance, beta_disp), mean),
            n_trajectories = first(n_trajectories),
            n_samples = first(n_samples),
            .groups = "drop") %>%
  group_by(group, compartment) %>%
  summarise(across(c(directionality, cumulative_distance, beta_disp),
                   list(mean = mean, sd = sd,
                        q025 = ~quantile(.x, 0.025), q975 = ~quantile(.x, 0.975)),
                   .names = "{.col}_{.fn}"),
            n_trajectories = mean(n_trajectories),
            n_samples = mean(n_samples),
            .groups = "drop")

Bac_stage.traj.summary

## Export
write.csv(Bac_stage.traj.all.df, "Bac_stage_traj_all_iterations.csv", row.names = FALSE)
write.csv(Bac_stage.traj.summary, "Bac_stage_traj_summary.csv", row.names = FALSE)

  ## Summary of all tables
summary_list <- list(
  ITS_Time  = ITS_time.traj.summary,
  ITS_Stage = ITS_stage.traj.summary,
  `16S_Time`  = Bac_time.traj.summary,
  `16S_Stage` = Bac_stage.traj.summary
)

all_summary <- bind_rows(summary_list, .id = "origin") %>%
  separate(origin, into = c("marker", "design"), sep = "_", remove = TRUE) %>%
  relocate(marker, design)

detail_list <- list(
  ITS_Time  = ITS_time.traj.all.df,
  ITS_Stage = ITS_stage.traj.all.df,
  `16S_Time`  = Bac_time.traj.all.df,
  `16S_Stage` = Bac_stage.traj.all.df
)

all_detail <- bind_rows(detail_list, .id = "origin") %>%
  separate(origin, into = c("marker", "design"), sep = "_", remove = TRUE) %>%
  relocate(marker, design)

## Last modifications
all_summary <- all_summary %>%
  mutate(treatment = coalesce(treatment, group)) %>%
  select(-group)

all_detail <- all_detail %>%
  mutate(treatment = coalesce(treatment, group)) %>%
  select(-group)

## Export
write.csv(all_summary, "all_sum.csv", row.names = FALSE) #Final Table
write.csv(all_detail,  "all_ite.csv", row.names = FALSE) #Table all Iterations
```

# Session info

```{r}         
sessionInfo()
```
